### 2024-09-23 analysis/paper/paper.Rmd
increase assessmentYrs to reduce impact of interannual variability

---
title: "Integration of national demographic-disturbance relationships into local caribou population viability projections can reduce uncertainty and inform monitoring decisions"
author:
  - name: Josie Hughes
    email: josie.hughes@ec.gc.ca
    institute: [ECCC]
    correspondence: true
  - name: Sarah Endicott
    email: sarah.endicott@ec.gc.ca
    institute: [ECCC]
    correspondence: false
  - name: Anna M Calvert
    email: anna.calvert@ec.gc.ca
    institute: [ECCC]
    correspondence: false
  - name: Cheryl A Johnson
    email: cheryl-ann.johnson@ec.gc.ca
    institute: [ECCC]
    correspondence: false

institute:
  - ECCC: Landscape Science and Technology Division, National Wildlife Research Center, Environment and Climate Change Canada, 1125 Colonel By Dr, Ottawa, ON, K1S 5B6, Canada
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      number_sections: no
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography: CaribouDemographicModel.bib
csl: "../templates/journal-of-forest-ecology-and-management.csl" # Insert path for the bib-style
abstract: |
  Across fifty-eight boreal caribou study areas in Canada, both survival and recruitment decrease with the percentage of the study area that is disturbed. There is variation in demographic rates among study areas, particularly where anthropogenic disturbance is low, but where disturbance is high we have no evidence of self-sustaining populations across the country. Demographic projections derived from local population-specific data are  uncertain for populations with limited monitoring. We propose a simple Bayesian population model that integrates prior information from national analysis of demographic-disturbance relationships with available local demographic data to reduce uncertainty in population viability projections, and reduce the risk that lack of local data will be used as a reason to delay conservation action. The model also acknowledges additional uncertainty and potential bias due to misidentification of sex or missing calves with a correction term derived from a simple model of the recruitment survey observation process.  We combine this Bayesian model with simulations of plausible population trajectories in a value of information analysis framework to show how the need for local monitoring varies with landscape condition, and to assess the ability of alternative monitoring scenarios to reduce the risk of errors in population viability projections. Where anthropogenic disturbance is high, reasonably accurate status projections can be made using only national demographic-disturbance relationships. At lower disturbance levels where uncertainty is higher, deploying fewer collars for more years may be better than intensive short-term monitoring for improving the reliability of status projections. 
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
editor_options: 
  chunk_output_type: inline
  markdown: 
    wrap: 72
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

<!-- With the following code you can access and display values from the yml header above. -->

<!-- Link to data here: https://drive.google.com/file/d/179lw6MPK4l3xGmded9kofzJkB6VlJVb9/view?usp=sharing -->

Keywords: `r rmarkdown::metadata$keywords`

Highlights: `r rmarkdown::metadata$highlights`

<!-- The following code chunk defines some general settings how code chunks should behave. -->

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figs/",
  dpi = 300
)
```

<!-- The actual document text starts here: -->

# Introduction

NOTE: please provide comments and suggestions for the introduction here:
<https://docs.google.com/document/d/1RD7bJ6yhtBJULjJFRJfHzObQscteyObM/edit?usp=sharing&ouid=111541730342945675365&rtpof=true&sd=true>.

The number of species threatened with extinction is high and getting
higher [@ipbes_global_2019]. Reliable information on the status of
populations and likely impacts of management are needed to inform
conservation action but effective monitoring is challenging and costly
[@legg_why_2006; @lindenmayer_effective_2018; @wintle_allocating_2010;
@dunham_strategic_2023]. Resources to support conservation and recovery
are limited [@buxton_balancing_2022; @mccarthy_financial_2012;
@gerber_conservation_2016; @wintle_spending_2019], delays increase risk
[@martin_acting_2012], and some populations have been monitored to
extinction [@lindenmayer_counting_2013]. A recent analysis of action
plans for Canadian species at risk found half of tasks are research and
monitoring, and this proportion is highest for species with higher risk
of extinction, highlighting a need to more carefully consider when,
where, and why additional information is needed to inform conservation
action [@buxton_balancing_2022]. Analysis of the implications of
monitoring outcomes for decisions can help clarify whether additional
monitoring is necessary to inform conservation action
[@bennett_when_2018; @dunham_strategic_2023; @runge_which_2011;
@gregory_structured_2012].

Analysis methods that can help make better use of available monitoring
data to inform decisions include Bayesian population modelling
[@schaub_integrated_2021] and value of information analyses
[@bennett_when_2018; @dunham_strategic_2023; @runge_which_2011;
@gregory_structured_2012]. To be useful, these methods must be tailored
to particular species and decision contexts. Here, we offer a simple
Bayesian population model for boreal caribou (*Rangifer tarandus
caribou*) that integrates information about the impacts of landscape 
change [@johnson_science_2020] with available demographic data [building on @eacker_web-based_2019] to reduce uncertainty in
population viability projections. Following the value of information analysis
framework proposed by Dunham et al. (2023), we combine this Bayesian
model with simulations of plausible population trajectories
[@stewart_climate-informed_2023; @dyson_existing_2022] to show how the
need for local monitoring varies with landscape condition, and to assess
the ability of alternative monitoring scenarios to reduce the risk of
errors in population viability projections. Our approach is tailored to
boreal caribou, but could be adapted for other species at risk found in
multiple populations that are costly to monitor and sensitive to habitat
disturbance.

Boreal caribou are widely and sparsely distributed through the Canadian
boreal region across a gradient of anthropogenic disturbance
[@eccc_scientific_2011; @johnson_science_2020] that includes some of the
most intact and inaccessible places remaining on earth
[@hirsh-pearson_canadas_2022; @ibisch_global_2016;
@venter_sixteen_2016]. Across 58 boreal caribou study areas, both
survival and recruitment decrease with the percentage of the study area
that is disturbed, and no populations of caribou are considered
self-sustaining in study areas with high anthropogenic disturbance
[@eccc_scientific_2011; @johnson_science_2020]. The causes of these
demographic-disturbance relationships include direct impacts on habitat
availability and use, and indirect impacts on predators and alternate
prey [@nbckc_boreal_2022; @eccc_scientific_2011]. Decreases in
recruitment as disturbance increases are also observed over time in
places where there is sufficient historic demographic data
[@rudolph_demographic_2017; @hervieux_widespread_2013], but many places
lack sufficient data to characterize local demographic-disturbance
relationships [@johnson_science_2020]. There is also variation in
demographic rates among study areas, particularly where anthropogenic
disturbance is low [@johnson_science_2020; @neufeld_primary_2021;
@fortin_forest_2017; @gagne_impact_2016], and this variation leads to
uncertainty in projections derived from national disturbance-demographic
relationships [@stewart_climate-informed_2023; @dyson_existing_2022].

Interventions such as wolf culls, maternal penning, and restoration of
linear features are intended to improve survival and reproduction of
populations that are not otherwise self-sustaining
[@mcnay_demographic_2022; @serrouya_predicting_2020], and there is a
need to assess and compare effects of boreal caribou conservation
actions with careful monitoring. Our focus in this paper is assessing
and projecting self-sustainability of populations in the absence of
these interventions, to clarify when additional information is needed to
reduce uncertainty about whether there is a need for action. Where
actions are taken, outcomes should be monitored; the simple model we
propose here does not include effects of management interventions, and
cannot be used to project outcomes for managed populations.

The most widely available source of boreal caribou demographic
information is from cows fitted with telemetry collars [e.g.
@courtois_aerial_2003; @rettie_dynamics_1998; @hervieux_widespread_2013;
@decesare_estimating_2012], though genetic methods are playing an
increasingly important role in some caribou monitoring programs [e.g.
@hettinga_estimating_2012; @mcfarlane_genetic_2018;
@moeller_integrating_2021; @flasko_fecal_2017; @jones_novel_2023;
@mcfarlane_identification_2022; @mcfarlane_spatial_2021]. Caribou cows
are typically captured via net-gun from aircraft in late winter, and
affixed with with collars that send a mortality signal when no longer
active [see @nbckc_nbckc_2021], thus allowing estimation of adult
survival (see @kaplan_nonparametric_1958 and @pollock_survival_1989 for
original approach, and @mcloughlin_declines_2003,
@hervieux_widespread_2013, @eacker_web-based_2019, @lamb_effectiveness_2024 and @dalgarno_bboutools_2024 for
examples of application to caribou). Recruitment is estimated from
aerial surveys of the composition of groups of animals that include
collared animals [@nbckc_nbckc_2022, @eacker_web-based_2019, @dalgarno_bboutools_2024]. Groups typically also contain
uncollared animals, and a critical assumption in these surveys is that
different demographic groups are equally detectable
[@ellington_divergent_2020]. Errors that can bias composition survey
results include misidentifying young bulls as cows or vice versa , and
failing to detect calves [@ellington_divergent_2020;
@decesare_estimating_2012].

Survival and recruitment survey data can be combined in a simple
recruitment-mortality population model [@hatter_moose_1991] that serves
as the basis for many herd viability assessments
[@decesare_estimating_2012; @serrouya_comparing_2017;
@eacker_web-based_2019; @eccc_scientific_2011; @johnson_science_2020; @dalgarno_bboutools_2024].
Various efforts have been made to improve caribou population viability
projections and growth rate estimates using Bayesian methods
[@moeller_integrating_2021; @mcnay_demographic_2022;
@eacker_web-based_2019; @lamb_effectiveness_2024; @dalgarno_bboutools_2024], following broader trends in the
analysis of population dynamics [@schaub_integrated_2021]. In Bayesian
population models, available demographic or population data (e.g.
abundance surveys, vital rate estimates, trend data) are combined into a
single analysis that can permit novel parameter estimation, and improve
precision, and account for multiple sources of uncertainty
[@schaub_integrated_2011; @eacker_assessing_2017;
@schaub_integrated_2021]. These models are proving broadly useful for
rapid detection of population declines, evaluation of management
actions, and viability analysis of critically endangered species
[@schaub_integrated_2021]. However, there is very little monitoring data
available for some caribou populations [@johnson_science_2020] and there
is a risk that lack of local data will be used as a reason to delay
conservation action [@boan_climate_2018]. Bayesian caribou population
models [@moeller_integrating_2021; @mcnay_demographic_2022;
@eacker_web-based_2019; @lamb_effectiveness_2024; @dalgarno_bboutools_2024] have not thus far included demographic-disturbance relationships, and we are concerned that models
informed only by local demographic data could be used to inflate
uncertainty about the status of poorly monitored populations.

We propose to reduce uncertainty in assessments and projections of local
caribou population viability by integrating prior information from
national analysis of demographic-disturbance relationships with
available local demographic data in a simple Bayesian population model
for boreal caribou. In our simple proof-of-concept model, information is 
shared among populations using priors derived from a published analysis of variation in demographic rates among populations across Canada [@johnson_science_2020]. A more
rigorous statistical approach would be to integrate multiple populations
into a hierarchical Bayesian analysis of variation within and among
populations [@leasure_hierarchical_2019]; we have not done this because
owners have not authorized the use of the data for this purpose,
obtaining permission to use caribou demographic data is difficult, and
most managers only have access to their own local data. Despite
limitations, our simple Bayesian model represents an important
improvement over less formally integrated methods for estimating and
projecting impacts of disturbance on population population viability
[@eccc_scientific_2011; @johnson_science_2020] because it reconciles and
accounts for uncertainties in local demographic data and
demographic-disturbance relationships. We show that integrating the most
informative predictor of outcomes for boreal caribou across their range
can substantially reduce uncertainty in local demographic projections
and status assessments, improve viability projections on changing
landscapes, and help clarify where, why and when additional data is
required to inform decisions and project outcomes on changing
landscapes. We hope that this approach can help clarify where available
information is sufficient for assessing and projecting population
viability, and thereby reduce the risk that uncertainty is used as a
reason to delay conservation action for imperiled populations.

While we believe that using available information to reduce uncertainty
is important, we also believe it is important to acknowledge remaining
uncertainty. In general, parametric Bayesian population models include a
number of simplifying assumptions, and misspecified models can yield
biased or overconfident results [@schaub_integrated_2021]. More
specifically, models that do not account for observation errors can be
misleading [@schaub_integrated_2021]. Although several authors have
pointed out that misclassification of adults and missing calves in
composition surveys can bias recruitment estimates from those surveys
[@ellington_divergent_2020; @decesare_estimating_2012] none have thus
far examined sensitivity of Bayesian caribou population models to these
sources of error [@moeller_integrating_2021; @mcnay_demographic_2022;
@eacker_web-based_2019; @lamb_effectiveness_2024; @dalgarno_bboutools_2024]. We address this gap with a bias correction term derived from a simple model of the recruitment survey observation
process.

# Methods

```{r setup2}
library(ggplot2)
library(dplyr)
library(stringr)
library(tidyr)
library(purrr)
library(ggpubr)
library(MetBrewer)
library(here)
library(tmap)
library(sf)

devtools::load_all(here::here())

#set paths
data_path_raw <- "analysis/data/raw_data/"
data_path_drvd <- "analysis/data/derived_data"

pal_nm <- "Redon"

fig_widths <- c(min = 1.18, single = 3.543, mid = 5.51, two = 7.48)

runScenarios = F #To do: set up to run full analysis from within this document.

resultScenario = "S10"
resultSet ="cmult6ay3aSf1"
figSet = c("diffs","distScns","EVSI","power","EVPIall")

if(runScenarios){
    source(here::here("analysis/scripts/0_nationalModelTrajectories.R"))
    source(here::here("analysis/scripts/0_bayesianModelExamples.R"))
    #create scenarios with sensitivity1.R and run in cloud
    #run viewSensitivityResults.R for S8 and S7
    #run EVSIPlots.R
}

for(f in figSet){
  if(f =="EVPIall"){addBit=""}else{addBit=resultSet}
  currentFile = here::here(paste0("figs/",resultScenario,"/",f,addBit,".png"))
  newFile = here::here(paste0("figs/forPaper/",f,addBit,".png"))
  file.copy(currentFile, newFile,overwrite=T)
}
```

## Caribou population monitoring data and composition survey bias correction

We focus on analyzing survival of collared females and the ratio of calves to cows in aerial
surveys of groups that include collared females [@nbckc_nbckc_2022],
because these remain the most widely available sources of demographic
information for boreal caribou populations [@johnson_science_2020].
Surveyed groups also contain uncollared animals so sample size is
typically larger for recruitment than for survival estimates. Errors
that can bias composition survey results include misidentifying young
bulls as cows or vice versa, and failing to detect calves
[@ellington_divergent_2020; @decesare_estimating_2012].

To derive a composition survey bias correction term we assume each group
of animals in a calf:cow composition survey contains one or more
collared adult females ($T$), and may also include: uncollared adult
females misidentified as young bulls or unknown sex ($U$); correctly
identified uncollared adult females ($V$); young bulls correctly
identified as male or unknown sex ($O$); young bulls misidentified as
uncollared adult females ($P$); observed calves ($J$); and unobserved
calves ($K$). The apparent number of adult females in the group is
$T+V+P=Tw$, where $w$ is the observed ratio of the apparent number of
uncollared adult females to collared animals in the recruitment survey.
The ratio of young bulls to uncollared adult females in the group is
$q = (P+O)/(U+V).$ Assuming an equal probability $u$ of misidentifying
young bulls as adult females and vice versa, we get $V=(U+V)(1-u)$ and
$P=(O+P)u$. Given a probability $z$ of missing calves, we get
$J=(J+K)(1-z)$. Given these assumptions and definitions, the sex and
bias-corrected recruitment rate $X$ can be written as a function of the
observed calf:cow ratio $R$, the cow multiplier $w$, the ratio of young
bulls to adult females $q$, and the misidentification probabilities $u$
and $z$ (see Supplement B for details). For simplicity, we combine the
probabilities into a bias correction term $c$ that is integrated into
the Bayesian population model and simulations (below):
$$c=\frac{w(1+qu-u)}{(w+qu-u)(1-z)}; X=cR/2$$

## Anthropogenic disturbance and monitoring scenarios

For this investigation we assume collars are deployed in January, and
last for up to 6 years. The target number of collared females $n$ varies
among scenarios, as does the total duration of the monitoring program
$d$ and the ratio of uncollared to collared adults in the recruitment
survey $w$ (Table \@ref(tab:monitoringScenarios)). Collars lost to
mortality may be replaced each year ($o=1$), or there may be a 3 year
gap between deployments ($o=4$).

We focus on exploring a set of scenarios in which buffered anthropogenic
disturbance, measured as the percentage of area within 500 m of some
type of anthropogenic disturbance [@johnson_science_2020;
@eccc_scientific_2011], increases by 1% per year (Figure
\@ref(fig:figDisturbance)). The duration of monitoring prior to the
projection period varies among scenarios from 1 to 24 years. The
projection period begins in 2024, and we consider how uncertainty
changes as the projection period increases from 5 to 20 years. The
amount of anthropogenic disturbance at the beginning of the projection
period varies from low (0%) through medium-low (20%) and medium-high
(40%) to high (60%). These simple hypothetical scenarios span a large
range to show how the need for local monitoring to reduce uncertainty
depends on landscape state, not to represent the reality of landscape
change. Scenarios in which anthropogenic disturbance remains the same or
decreases over time are included in Supplement D.

```{r monitoringScenarios, echo=FALSE,message=FALSE}

source(here::here("analysis/scripts/formatScenarios.R"))

knitr::kable(tbl,caption="Monitoring scenario parameters.")
```

```{r figDisturbance, fig.cap="Anthropogenic disturbance and monitoring scenarios. Anthropogenic disturbance increases by 1% per year, and the amount of disturbance at the beginning of the projection period varies from low (0%), through low-med (20%) and med-high (40%) to high (60%). These simple hypothetical scenarios are intended to show how the need for local monitoring to reduce uncertainty varies with landscape state, and not intended to reflect real landscape change. In all scenarios, the projection period begins in 2024, and the monitoring period occurs prior to the projection period.", cache=TRUE}
knitr::include_graphics(here::here("figs/forPaper/distScnscmult6ay3aSf1.png"))

```

```{=html}
<!--
*Monitoring WG input from BC, ON, NL, (MB pending): lots of detail provided but generally those numbers suggest we could try values of ~3-10 [e.g. 3,6,9?] for the cow multiplier, 5-50 for collars [e.g. 5,10,20,40,60?], [not sure about renewal interval, but see SaraM email for some insights]. We know (e.g. from the MonitoringWG perspectives report/update to Rettie summary) that number of years of monitoring can be as low as 2 or as many as 20+

*from Anna: withouot access to actual estimates of R and S for different ranges, some potentially interesting other ranges to compare to could be an Alberta range or two, where there are ++years of R and S estimates but really low lambda (E.g. ESAR and WSAR have 27+ yrs consecutive monitoring but declining numbers and high disturbance), some sections of NWT1 that are thought to be self-sustaining and have pretty good data sets (e.g. Dehcho North and South, Hay River Lowlands with 13+yrs of S and R) -could also use other MB ranges as examples for fairly sparse data (though not as bad as ON), and QC or LAB ranges as examples of good data (e.g. how about Charlevoix for a very well monitored but tiny pop, northern QC for really broad-expanse pop, or Labrador pops have long term monitoring data but uncertain status...)

*feedback from MonitoringWG contacts includes details on collar numbers for BC, ON, NL, (MB pending); e.g. in Ontario they aim for 20-30 collars in smaller pops, 35+ in bigger pops; Labrador aims to have between 5-25 collars monitored at any one time, depending on pop size; in BC, real numbers of collared animals have ranged been between 3-59 per pop/year, varying both across-yrs-within-pop and across-pop-within-yr [with COVID yrs by far the lowest].

relevant refs: [@rettie_summary_2017];?

TO DO: Fire scenarios? Add high fire scenario in appendix of one by one sensitivity checks.

TO DO: run low initial population size scenario, put in appendix of one by one sensitivity checks? 
-->
```
## Simulation of population dynamics and monitoring

To investigate the effectiveness of various monitoring strategies for
reducing uncertainty about population viability, and to show how this
varies with landscape condition, we begin with a model for simulating
example population dynamics and observations of those dynamics. These
example population trajectories are informed by prior information about
the distribution of possible parameter values and outcomes from the
national analysis of demographic-disturbance relationships
[@johnson_science_2020] so we call it the "national" model. Our method
for selecting parameter values for each example trajectory yields a set
of alternative models (hypotheses) about the state of a particular
caribou population; these alternative scenarios informed by prior
beliefs are used to assess the expected value of alternative sampling
strategies [@runge_which_2011; @dunham_strategic_2023]. We also use the
simulated trajectories to confirm that the Bayesian model successfully
integrates local data with priors from the national model. For
reference, all variables and parameters are defined in supplemental
Table A1.

### A model for simulating example population dynamics informed by national disturbance-demographic relationships

To simulate demographic trajectories for example populations we used a
modified version of Johnson et al's [-@johnson_science_2020] demographic
model described by Dyson et al. [-@dyson_existing_2022]. Given a
population of post-juvenile females at the beginning of year $t$,
$\dot{N}_t$, the number that survive to the census, $\dot{W}_t$, is
binomially distributed with true survival probability $\dot{S}_t$:
$\dot{W}_{t} \sim \text{Binomial}(\dot{N}_t,\dot{S}_t)$. The calf:cow
ratio $\dot{R}_t$ is adjusted for sex ratio, misidentification biases
(Supplement B) and (optionally) a delayed reproduction term to account
for the reality that yearlings generally do not breed
[@eacker_web-based_2019; @decesare_estimating_2012] to get maximum
potential recruitment of female calves per cow $\dot{X}_t$:

\begin{equation}
  \dot{X}_t=\frac{\dot{c}\dot{R}_t/2}{1+\dot{c}\dot{R}_t/2};\dot{c}=\frac{w(1+\dot{q}\dot{u}-\dot{u})}{(w+\dot{q}\dot{u}-\dot{u})(1-\dot{z})},
  (\#eq:dotX)
\end{equation}

where $w$ is the ratio of observed adult females to
collared animals, $\dot{q}$ is the ratio of young bulls to adult
females, $\dot{u}$ is the probability of misidentifying young bulls as
adult females and vice versa, and $\dot{z}$ is the probability of
missing a calf. We examine sensitivity to these unknown parameters by
selecting values for each simulated example population from plausible
ranges: $\dot{u}$ and $\dot{v}$ vary uniformly between 0 and 0.2, and
$\dot{q}$ varies uniformly between 0 and 0.6.

Realized recruitment rate varies with population density
[@lacy_vortex_2017], and the number of juveniles recruiting to the
post-juvenile class at the census $(\dot{J}_{t})$ is a binomially
distributed function of the number of surviving post-juvenile females
and the adjusted recruitment rate:

\begin{equation}
  \dot{J}_{t} \sim \text{Binomial}(\dot{W}_t,\dot{X}_t[p_0-(p_0-p_k)(\frac{\dot{W}_t}{k})^b]\frac{\dot{W}_t}{\dot{W}_t+a}).
  (\#eq:dotJ)
\end{equation}

Given the values of parameters $p_0,p_k,a,$ and $b$ used by Johnson et
al. (2020) (Table A1), recruitment rate is lowest $(0.5\dot{X}_t)$ when
$\dot{N}_t=1$, approaches a maximum of $\dot{X}_t$ at intermediate
population sizes, and declines to $0.6\dot{X}_t$ as the population
reaches carrying capacity of $k=50000$.
The post-juvenile female population in the next year includes both
survivors and new recruits:

\begin{equation}
\dot{N}_{t+1}=\text{min}(\dot{W}_t+\dot{J}_t,r_{max}\dot{N}_t).
  (\#eq:dotN)
\end{equation}

Interannual variation in survival and recruitment is modelled using
truncated beta distributions [rtrunc function;
@novomestky_package_2016]:
$\dot{R}_t \sim \text{TruncatedBeta}(\bar{R}_t,v_R,l_R,h_R); \dot{S}_t \sim \text{TruncatedBeta}(\bar{S}_t,v_S,l_S,h_S)$.
Coefficients of variation among years $(v_R,v_S)$ and
maximum/minimum values $l_R,h_R,l_S,h_S$ for recruitment and survival
are given in Table A1. Expected recruitment ($\bar{R}_t$) and survival
($\bar{S}_t$) vary with disturbance according the beta regression models
estimated by Johnson et al. [-@johnson_science_2020]:

\begin{equation}
\bar{R}_t \sim Beta(\mu^R_t,\phi^R); 
log(\mu^R_t) = \dot{\beta}^R_0+\dot{\beta}^R_a A_t+\dot{\beta}^R_f F_t,
  (\#eq:barR)
\end{equation}

\begin{equation}
\bar{S}_t \sim (46 \times Beta(\mu^S_t,\phi^S)-0.5)/45;
log(\mu^S_t) = \dot{\beta}^S_0+\dot{\beta}^S_a A_t.
  (\#eq:barS)
\end{equation}

$\phi^R \sim \text{Normal}(19.862,2.229)$ and
$\phi^S \sim \text{Normal}(63.733,8.311)$ are precisions of the Beta
distributed errors [@ferrari_beta_2004]. At the beginning of a
simulation for an example population, regression coefficient values are
sampled from Gaussian distributions (see Table \@ref(tab:priors)) and
the population is assigned to quantiles of the Beta error distributions
for survival and recruitment. The population remains in these quantiles
as disturbance changes over time, so there is substantial persistent
variation in recruitment and survival among example populations (Figure
\@ref(fig:exampleTrajectories)). To ensure results are comparable across
monitoring scenarios and that populations remain extant through the
monitoring period, all simulations are started with $N_0=5000$ animals
24 years prior to the start of the projection period.

```{r exampleTrajectories, fig.cap="Examples of simulated population trajectories from a model informed only by national disturbance-demographic relationships with (top row) and without (bottom row) delayed reproduction. To highlight how variation among trajectories persists as disturbance changes we have reduced interannual variation by 75\\% in these examples ($v_R=0.115,v_S=0.02175$). Orange dots show means of 500 example trajectories. For reference, black lines and and shaded bands show means and 95\\% predictive intervals from Johnson et al's (2020) regression models without adjustment for misidentification bias and delayed reproduction.", cache=TRUE}

knitr::include_graphics(here::here("analysis/paper/figs/DemographicRatesDelay.png"))

```

### Simulation of observed survival and recruitment

In our simulated example populations, mortality risk does not vary over
time or among individuals, so the observed state (dead or alive) of a
collared animal $i$ in month $m$ and year $t$ depends on the true
survival rate $\dot{S}_t$ as follows: 
\begin{equation}
\hat{I}_{i,t,m} \sim \text{Bernoulli}(\dot{S}^{1/12}_t\hat{I}_{i,t,m-1}).
  (\#eq:hatI)
\end{equation}

The number of cows in the composition survey $\hat{W}_t$ is given by the
number of collared cows $\hat{T}_t$ and the apparent number of adult
females per collared female observed in the composition survey $w$:
\begin{equation}
\hat{W}_t = w\hat{T}_t.
  (\#eq:hatW)
\end{equation}
The number of observed calves $\hat{J}_t$ also depends on
the unadjusted apparent recruitment for the population $\dot{R}_t$:
\begin{equation}
\hat{J}_t \sim \text{Binomial}(\hat{W}_t,\dot{R}_t).
  (\#eq:hatJ)
\end{equation}

## Bayesian integration of local demographic data and national disturbance-demographic relationships

Local demographic and covariate data provided as inputs to the Bayesian
population model include the observed state (dead or alive) of each
collared animal in each month ($\hat{I}_{i,t,m}$), the observed number
of adult females and calves in the recruitment surveys
($\hat{W}_t, \hat{J}_t$), the apparent number of adult females per
collared female in the composition survey $w$, and the measures of
anthropogenic disturbance and fire ($A_t, F_t$). Recruitment rate $R_t$
is estimated from the observed number of calves $\hat{J}_t$ and adult
female caribou $\hat{W}_t$ assuming a Binomial distribution:
\begin{equation}
\hat{J}_t \sim \text{Binomial}(\hat{W}_t,R_t).
  (\#eq:hatJEst)
\end{equation}

Where appropriate, the Bayesian population model and national
demographic model share the same structure, and we have selected
notation to highlight the similarities. The difference between equations
\@ref(eq:hatJ) and \@ref(eq:hatJEst) is that in the simulation model an
unadjusted apparent recruitment rate $\dot{R}_t$ is selected and used to
simulate observations. In the Bayesian model, in contrast, observations
are provided, and used to estimate the unobserved recruitment rate
$R_t$.

In both models recruitment probability is a function of anthropogenic
disturbance and fire with a log link and Beta distributed interannual variation [@johnson_science_2020;
@stewart_climate-informed_2023; @dyson_existing_2022]. The coefficient of variation $\nu_R$ is constant, while the mean $\tilde{R_t}$ and standard deviation $\sigma_{R,t}$ vary with disturbance:
\begin{equation}
log(\tilde{R_t})=\beta^R_0+\beta^R_a A_t+\beta^R_f F_t
  (\#eq:Rtilde)
\end{equation}

\begin{equation}
R_t \sim \text{Beta}(\tilde{R_t},\sigma_{R,t});
\sigma_{R,t} = min(\nu_R \tilde{R_t},\sqrt{\tilde{R_t} (1-\tilde{R_t})})
(\#eq:R)
\end{equation}

In the simulation model (eq \@ref(eq:barR)), regression coefficient
values ($\dot{\beta}^R_0,\dot{\beta}^R_a,\dot{\beta}^R_f$) are selected
for each example population and used to calculate recruitment rate. In
the Bayesian model (eq \@ref(eq:R)), unobserved regression coefficients
($\beta^R_0,\beta^R_a,\beta^R_f$) are estimated from local recruitment
survey and disturbance observations.

To account for bias and additional uncertainty caused by unknown
misidentification biases (Supplement B) and delayed age of age of first
reproduction [@eacker_web-based_2019; @decesare_estimating_2012] the
estimated recruitment probability is adjusted to get expected female
recruits per female: 
\begin{equation}
X_t=\frac{cR_t/2}{1+cR_t/2}; c=\frac{w(1+qu-u)}{(w+qu-u)(1-z)}.
  (\#eq:X)
\end{equation}

The composition survey bias correction term $c$ depends on the known
apparent number of adult females per collared animal $w$, the unknown
ratio of young bulls to adult females $q$, the unknown probability of
misidentifying young bulls as adult females and vice versa $u$, and the
unknown probability of missing a calf $z$ (Supplement B). Prior
uncertainty about the value of the bias correction term $c$ is
approximated with a Log-normal distribution (Supplement B). Given the
apparent number of adult females per collared animal $w$ in each
monitoring scenario (Table \@ref(tab:monitoringScenarios)), we calculate
the mean and standard deviation of 10,000 samples of $\log{ c}$ assuming
the ratio of young bulls to adult females $q$ varies uniformly between 0
and 0.6, the adult misidentification probability $u$ varies uniformly
between 0 and 0.2, and the probability of missing a calf $z$ varies
uniformly between 0 and 0.2. To maintain consistency with Johnson et al.
[-@johnson_science_2020] we also provide an option to only adjust for
sex ratio and misidentification biases $X_t=cR_t/2$.

Following others [@eacker_web-based_2019; @lamb_effectiveness_2024], Kaplan-Meier
estimates of observed adult female survival probability $\hat{S}_t$ and
associated precisions $\tau_t$ are estimated from known-fate radio
collar data using the *survival* R package [@therneau_modeling_2000],
with precisions for years with no observed mortality estimated using a
separate Bayesian model. Observed estimated survival is modelled as a
Gaussian distributed function of the true survival probability,
truncated between 0 and 1 [@eacker_web-based_2019,@lamb_effectiveness_2024]:
\begin{equation}
\hat{S_t} \sim \text{Normal}(S_t,1/\tau_t)T(0,1).
  (\#eq:hatSEst)
\end{equation}

For consistency with Johnson et al. [-@johnson_science_2020], we model survival
probability as an adjusted function of buffered anthropogenic
disturbance $A_t$ with a log link and Beta distributed interannual variation with coefficient of variation $\nu_S$, mean $\tilde{S_t}$ and standard deviation $\sigma_{S,t}$ [@stewart_climate-informed_2023;
@dyson_existing_2022]: 
\begin{equation}
\tilde{S_t}=(46 e^{\beta^S_0+\beta^S_a A_t}-0.5)/45
  (\#eq:Stilde)
\end{equation}

\begin{equation}
S_t \sim \text{Beta}(\tilde{S_t},\sigma_{S,t});
\sigma_{S,t} = min(\nu_S \tilde{S_t},\sqrt{\tilde{S_t} (1-\tilde{S_t})})
(\#eq:S)
\end{equation}

To assess sensitivity of results to the survival estimation method and
handle cases in which there is little or no survival data, we also
implemented an alternative parametric exponential survival model,
wherein the observed state (alive or dead, $\hat{I}_{i,t,m}$) of an
individual animal $i$ in year $t$ and month $m$ is Bernoulli distributed: 
\begin{equation}
\hat{I}_{i,t,m} \sim \text{Bernoulli}(S^{1/12}_t\hat{I}_{i,t,m-1}).
  (\#eq:hatIEst)
\end{equation}

All regression coefficients are assumed to be Gaussian distributed, with
prior means from Johnson et al's (2020) regression model estimates
(Table \@ref(tab:priors)). We calibrated the standard deviations of the
priors so that the the 95% prior predictive intervals for survival and
recruitment from the Bayesian model match the 95% predictive intervals
from Johnson et al's regression model (Supplement C, Figure
\@ref(fig:examplePredictions)).

```{r priors, echo=FALSE,message=FALSE}

source(here::here("analysis/scripts/formatPriors.R"))

knitr::kable(tbl,caption="Prior means and standard deviations of Gaussian distributed survival and reproduction model parameters. Standard deviations are calibrated so that the 95\\% prior predictive intervals for survival and recruitment from the Bayesian model match the 95\\% predictive intervals from Johnson et al's regression model (Supplement C).")
```

As in the simulations, we assume a simple demographic model, with a
census in late winter of surviving females and new recruits to the
post-juvenile (aged 1+) class. The number of post-juvenile females that
survive from year $t$ to the census $W_t$ is binomially distributed with
survival probability $S_t$: $W_{t} \sim \text{Binomial}(N_t,S_t)$. The
number of juveniles recruiting to the post-juvenile class at the census
is a Poisson distributed function of the number of surviving
post-juvenile females and the adjusted recruitment rate $X_t$:
$J_{t} \sim \text{Poisson}(W_tX_t)$. The post-juvenile female population
in the next year includes both survivors and new recruits:
$N_{t+1}=W_t+J_t$. Annual population growth rate $\lambda_t$ is set to 0
when $N_t$ is 0, and is otherwise $\lambda_t=N_{t+1}/N_t$. For more
stable estimates of population status, we report average population
growth over a period of $y=3$ years, calculated as
$\bar{\lambda}_{y,t}=\sum_{x=t-y+1}^{x=t}\lambda_x$. The initial
post-juvenile population size $N_0$ is 5000 animals. For reference, all
variables and parameters are defined in supplemental Table A1.

```{=html}
<!-- Annotated equations for poster.
$$J_{t} \sim \text{Poisson}(W_tX_t)$$
$$\underbrace{N_{t+1}}_{\text{Adult females in year t+1}}=\underbrace{W_t}_{\text{Surviving adults}}+\underbrace{J_t}_{\text{New recruits to adult age class}}$$
$$W_{t} \sim \text{Binomial}(N_t,\underbrace{S_t}_{\text{Adult survival probability}})$$
$$S_t=\underbrace{\text{min}(46 \tilde{S_t}-0.5)/45,1); \log(\tilde{S_t})}_{\text{Adjustment as in in Johnson et al. 2020}}=\beta^S_0+ \underbrace{\beta^S_aA_t}_{\text{Anthropogenic disturbance effect}}+\underbrace{\epsilon^S_t}_{\text{Random effect of year}}.$$
$$J_{t} \sim \text{Poisson}(W_t\underbrace{R_t/2}_{\text{Recruitment rate}})$$
$$\log(R_t)=\beta^R_0+\underbrace{\beta^R_a A_t}_{\text{Anthropogenic disturbance effect}}+ \underbrace{\beta^R_fF_t}_{\text{Fire effect}}+\underbrace{\epsilon^R_t}_{\text{Random effect of year}}.$$
$$\underbrace{\hat{C}_t}_{\text{Observed calves}} \sim \text{Binomial}(\underbrace{\hat{W}_t}_{\text{Observed cows}},R_t)$$
$$\underbrace{\hat{S_t}}_{\text{Observed survival (Kaplan-Meier estimate)}} \sim \text{Normal}(S_t,1/\tau_t).$$ -->
```
```{=html}
<!--
Gritty details for the record: \* note distribution of random effect parameters are truncated to
ensure R and S remain within reasonable bounds (not guaranteed given log
link). Not yet documented here. \* note Eacker survival method gives biased results if there is left censoring, or months without animals at risk. To address this we drop left censored observations and only estimate survival when at least some animals remain at risk by month 12.
\*
see alternative choices made in this example
<https://poissonconsulting.shinyapps.io/bboushiny/>. See also Lamb et al, who uses two step KM estimation process for survival like Eacker. Stick with published examples. \*
-->
```
Models and analysis were implemented using the R language
[@r_core_team_r_2023] and a variety of *tidyverse* packages
[@wickham_welcome_2019]. We use the open source Gibbs sampling software
JAGS [@plummer_jags_2023] and the *R2jags* R package [@su_r2jags_2021]
to fit the Bayesian model. We ran 15,000 iterations of 4 chain,
discarded the first 10,000 interations, and used a thinning rate of 2.
We assessed convergence for a small random subset of example
trajectories using a threshold for the Gelman-Rubin diagnostic
[@gelman_inference_1992] of r-hat \< 1.1, and visually inspected
posterior autocorrelation and traceplots produced by the *mcmcplots* R
package [@curtis_mcmcplots_2018]. Code required to run analyses and
reproduce figures in this paper is available at
<https://github.com/LandSciTech/Caribou-Demographic-Projection-Paper>,
and in the *caribouMetrics* R package available on GitHub at
<https://github.com/LandSciTech/caribouMetrics>.

## Probability of correct status assessment and expected value of information

In order to assess whether a monitoring program is likely sufficient to
inform decisions, it is important to select outcome measures that are
relevant to decision making [@gregory_structured_2012]. For boreal
caribou and other endangered species, a relevant outcome is the
probability of correctly predicting the true population status
(increasing or decreasing) [@dunham_strategic_2023]. To calculate this
outcome for each disturbance and monitoring scenario (Fig
\@ref(fig:figDisturbance)) we fit the Bayesian model to simulated
observations from 500 example population trajectories (e.g. Fig
\@ref(fig:exampleTrajectories)), and calculated the percentage of these
500 examples in which the Bayesian posterior mean population growth rate
and the true population growth rate over the full projection period are both greater than 0.99, or both less than or equal to 0.99.

As suggested by Dunham et al. [-@dunham_strategic_2023] we also
calculate the expected value of perfect information (EVPI) and the
expected value of sample information (EVSI) [@runge_which_2011;
@yokota_value_2004]. We assign a value or utility of
$\eta_{\alpha,\psi}=1$ when the true state of the population $\psi$
matches the assigned status $\alpha$, and value of $\eta_{\alpha,\psi}=0$
when the status assessment is incorrect. This is "the value ascribed by
the decision-maker to the outcome, and is thus the measure of management
performance" [@runge_which_2011]. The expected value of current
information is the average utility of decisions informed by knowledge of
the disturbance scenario $\tau$ (Figure \@ref(fig:figDisturbance)) and
priors from the national model:
\begin{equation}
EV_{current} = max_{\alpha}\sum_\psi[P(\psi|\tau)\eta_{\alpha,\psi}],
  (\#eq:EVcurrent)
\end{equation}
where $P(\psi|\tau)$ is the prior probability that the
true state of the population is $\psi$, and each population is assigned
a status $\alpha$ informed by the prior. The expected value under
certainty (i.e. the expected value of a decision when the true state of
the population is known) is $EV_{certainty}=1$. The difference between
$EV_{certainty}$ and $EV_{current}$ gives EVPI, a measure of the maximum
possible benefit that could theoretically be gained if the system were
perfectly known.

EVSI measures expected improvement from a particular sampling strategy
as the difference between the expected value with and without additional
sample information: 
\begin{equation}
EVSI_{\tau} = E_x[max_{\alpha}\sum_\psi[P(\psi|x,\tau)\eta_{\alpha,\psi}]] - EV_{current},
  (\#eq:EVSI)
\end{equation}
where $x$ is the sample information from a particular
example trajectory. In this study, we treat simulated observations from
500 example trajectories for each disturbance and monitoring scenario
$\tau$ (Figure \@ref(fig:figDisturbance)) as samples from the
distribution of possible values of $x$, and assume each of these sample
results is equally probable. $P(\psi|x,\tau)$ is the posterior
probability that the true state of the population is $\psi$, and each
population is assigned a status $\alpha$ informed by the sample
information.

To aid in interpretation of results and to highlight the advantages of
selecting a metric that is directly relevant to the management objective
of stable or increasing populations, we also show how integrating local
monitoring data alters the distribution of differences between true
population growth rates and Bayesian posterior mean estimated growth
rates.

# Results

Analysis of simulated example demographic trajectories confirms that our
Bayesian model can effectively combine local demographic data with
priors from national demographic-disturbance relationships to reduce
uncertainty in boreal caribou demographic projections (Figures
\@ref(fig:examplePredictions) and \@ref(fig:figDiffs)). As expected,
uncertainty about the population growth rate decreases as sampling
effort in each year (i.e. number of collars) or the number of years of
monitoring (Figure \@ref(fig:figDiffs)) increase. Uncertainty is higher
when the forecast horizon is long (compare 5 and 20 year projections in
Figure \@ref(fig:figDiffs)), and when interannual variation is high (Figure
\@ref(fig:examplePredictions)). Uncertainty is also higher in our high
anthropogenic disturbance scenarios (Figure \@ref(fig:figDiffs)) because
stochastic variation in population growth rates increase as populations
become very small (as in top right panel of Figure
\@ref(fig:examplePredictions)).

Our analysis becomes more informative when we examine the probability of
correctly assessing population status (Figure \@ref(fig:figStatusErrors))
and the expected value of information (Figures \@ref(fig:figEVSI) and E.1).
When anthropogenic disturbance is high (right panels of Figures
\@ref(fig:figStatusErrors), \@ref(fig:figEVSI), and E.1)), projections
of population status informed only by priors from the national
demographic-disturbance relationships are accurate, and adding
additional local monitoring data does not improve accuracy.
Interestingly, when the projection period is long and anthropogenic
disturbance is high, we see a negative EVSI (lower right panel of Figure
\@ref(fig:figEVSI)), and we do not see an associated decrease in the
probability of correctly assessing population status (lower right panel
of Figure \@ref(fig:figStatusErrors)). This means that the long-term
projections informed by local monitoring data are on average less
precise but not less accurate, which can happen when there is very small
prior probability that population is viable (orange band in the top
right panel of Figure \@ref(fig:examplePredictions) falls below 1). At
lower levels of disturbance, local monitoring can reduce uncertainty,
but in some circumstances there are diminishing returns as the number of collars deployed or
the number of years of monitoring increases (Figures
\@ref(fig:figStatusErrors) and \@ref(fig:figEVSI)). High interannual variation makes reducing uncertainty more difficult (Figures
\@ref(fig:figStatusErrors) and \@ref(fig:figEVSI)).The shapes of the
relationships between monitoring effort and the probability of correct
status assessment depend on how anthropogenic disturbance is changing
over time, but there are diminishing benefits of increased monitoring
regardless of whether disturbance is increasing or decreasing
(Supplement D). Adding unknown bias increases uncertainty, but these
increases are not large for the set of parameters we investigate
(Supplement F).

```{r examplePredictions, fig.cap="Prior (a) and posterior predictions from the Bayesian model for population trajectories with high (b: $v_R=0.46,v_S=0.087$) and lower (c: $v_R=0.23,v_S=0.0435$) interannual variation. In this example scenario anthropogenic disturbance increases by 2\\% per year from 0 to 100\\%. The population is monitored for 16 years with 30 collars per year, and 6 cows per collared cow in the composition surveys. The prior means and 95\\% predictive intervals (top row, orange lines and shading) are similar to the means and ranges between the 2.5\\% and 97.5\\% quantiles of 3000 simulated survival and recruitment trajectories from the national model (top row, blue lines and shading). Local population data (open triangles) reduces uncertainty more when interannual variation is lower (bottom row), but when disturbance is very high local data is not required to predict viability (population growth <= 0.99). When projected populations become very small, stochastic variation in projected population growth rate increases (as in the top right panel).", cache=TRUE}

knitr::include_graphics(here::here("analysis/paper/figs/bayesianModelExamples.png"))
```

```{r figDiffs, fig.cap="The distribution of differences between true population growth rates and Bayesian posterior mean growth rates varies with the amount of anthropogenic disturbance (columns), the number of years projected (rows), and monitoring effort (see Figure \\@ref(fig:figDisturbance) for scenarios). Differences decrease as the number of years of monitoring and the number of collars increase. Differences are larger when anthropogenic disturbance is high because stochastic variation in population growth rate is higher for small populations. In these examples the composition survey cow multiplier $w$ is 6, collars are renewed each year, and interannual variation is lower ($v_R=0.23,v_S=0.0435$).", cache=TRUE}
knitr::include_graphics(here::here("figs/forPaper/diffscmult6ay3aSf1.png"))
```

```{r figStatusErrors, fig.cap="The probability of a correct status assessment varies with the amount of anthropogenic disturbance (columns), the number of years projected (rows), the amount of interannual variation (high: $v_R=0.46,v_S=0.087$, low: $v_R=0.23,v_S=0.0435$), and monitoring effort (see Figure \\@ref(fig:figDisturbance) for scenarios). When disturbance is high the probability of correct status assessment is high even without local monitoring. Where disturbance is low or moderate, increasing the number of years of monitoring and the number of collars reduces uncertainty, but there are diminishing returns on additional monitoring (number of years or number of collars) as the number of years or collars increases. In these examples the composition cow multiplier $w$ is 6 and collars are renewed each year; varying these parameters does not alter qualitative conclusions (Supplements D and E).", cache=TRUE}
knitr::include_graphics(here::here("figs/forPaper/powercmult6ay3aSf1.png"))
```

```{r figEVSI, fig.cap="The expected value of sample information (EVSI) varies with the amount of anthropogenic disturbance (columns), the number of years projected (rows), the amount of interannual variation (high: $v_R=0.46,v_S=0.087$, low: $v_R=0.23,v_S=0.0435$), and monitoring effort. When disturbance is high EVSI is low, indicating that local sampling does not alter status assessments. Where disturbance is lower, EVSI increases with the number of years of monitoring and the number of collars, but in some circumstances there are diminishing returns on additional monitoring (number of years or number of collars) as the number of years or collars increases. See Figure \\@ref(fig:figStatusErrors) for additional scenario details.", cache=TRUE}

knitr::include_graphics(here::here("figs/forPaper/EVSIcmult6ay3aSf1.png"))
```

# Discussion

NOTE: Please provide comments and suggestions for the discussion here:
<https://docs.google.com/document/d/1RD7bJ6yhtBJULjJFRJfHzObQscteyObM/edit>.

## Bayesian integration of local demographic data & national demographic-disturbance relationships

The availability of boreal caribou demographic data varies widely across
Canada due to large variation in habitat condition, population
condition, available resources for monitoring, and remoteness. In a few
regions enough data has been collected over time to characterize local
demographic-disturbance relationships [@rudolph_demographic_2017;
@hervieux_widespread_2013], but most populations have only been sparsely
monitored for snapshot estimates of trend or abundance
[@johnson_science_2020]. As a result, in most regions it is necessary to
make assumptions informed by observed variation in demographic rates
with disturbance among populations and other relevant information about
impacts of disturbance on caribou. However, high observed variation in
demography where anthropogenic disturbance is low leads to high
uncertainty in projections informed by national demographic-disturbance
relationships alone [@stewart_climate-informed_2023;
@dyson_existing_2022]. In some cases, predictions of population growth
rate informed by national demographic-disturbance relationships appear
to disagree with estimates from local demographic data
[@eccc_scientific_2011; @dyson_existing_2022], but without formal
methods of accounting for uncertainty it can be difficult to distinguish
true discrepancies from stochastic variation.

We developed a simple Bayesian population model [building on
@eacker_web-based_2019] to help address these challenges by
acknowledging uncertainty associated with local demographic data and
national demographic-disturbance relationships, and integrating these
two sources of information to reduce uncertainty in projections of the
impacts of changing landscape conditions. This Bayesian approach
represents an important improvement over less formally integrated
methods for estimating and projecting impacts of disturbance on
population population viability [@eccc_scientific_2011;
@johnson_science_2020] because it reconciles and accounts for
uncertainties in local demographic data and demographic-disturbance
relationships.

Our model assumes that increasing cumulative anthropogenic disturbance
causes a log-linear decrease in caribou demographic rates (equations 4,
5, 10 and 13), reflecting a scientific consensus informed by a number of
lines of evidence that cumulative anthropogenic disturbance has a
negative impact on boreal caribou [@eccc_scientific_2011;
@johnson_science_2020; @gouvernement_du_quebec_revue_2021;
@rudolph_demographic_2017; @sorensen_determining_2008;
@hervieux_widespread_2013]. Local data can be integrated to estimate
local demographic-disturbance parameters that differ from the national
means, and in the absence of local data, appropriate priors lead to a
range of outcomes consistent with the observed range of variation across
the country [Figures 3 and 5 in @johnson_science_2020]. Others
[@moeller_integrating_2021; @mcnay_demographic_2022;
@eacker_web-based_2019; @dalgarno_bboutools_2024] have also used Bayesian models to estimate
caribou demographic rates from local monitoring data, but our addition
of prior information from national demographic-disturbance relationships
is novel, and important for reducing the risk that lack of local data
will be used as a reason to delay conservation action for highly
disturbed and poorly monitored populations. Integrating
demographic-disturbance relationships into a viability projection model
clarifies that no additional information is needed to project outcomes
where disturbance is high and actions to support recovery are not taken.

There are a variety of possible actions intended to improve survival and
reproduction of populations that are not otherwise self-sustaining
[@mcnay_demographic_2022; @serrouya_predicting_2020]. The simple models
in this paper do not include these actions, and cannot be used to
project outcomes for managed populations. Our intention is to reduce
uncertainty that can be used to delay conservation action; there is
nothing in our arguments, models or results that can or should be used
to discourage much needed monitoring of the effects of conservation
actions. Even when no additional monitoring is required for assessing
self-sustainability, there are a variety of reasons to continue
monitoring highly disturbed boreal caribou populations. The point we wish to emphasize is
that action is also required, and lack of monitoring should not be used
as a reason to delay conservation measures for highly disturbed
populations.

Variation in anthropogenic disturbance across the Canadian boreal region
is partially correlated with variation in climate and primary
productivity [@neilson_direct_2022] because forestry occurs in higher
productivity areas, and petroleum extraction activities are concentrated
in Western Canada. Effects of disturbance on vegetation and foodweb
dynamics vary with primary productivity [@neufeld_primary_2021], and
some have expressed concern about confounding of demographic-disturbance
relationships [@wilson_causal_2021]. One option for addressing these
concerns would be to develop regional demographic-disturbance
relationships for populations that are ecologically similar to one
another (e.g. caribou inhabiting forests that are sufficiently
productive for industrial forestry in Eastern Canada)
[@eccc_scientific_2024], and use the priors from regional
demographic-disturbance relationships to inform demographic projections
within the region. However, while acknowledging and allowing for
regional variation, we also need to guard against risks that uncertainty
will be inflated and used as an excuse to delay action, a tactic that
has been used in this and other decision contexts
[@bissonnette_controverses_2020; @boan_climate_2018]. In order to be
informative and credible, regional demographic-disturbance analyses must
include a sufficient number of observations spanning a range from high
to low anthropogenic disturbance. Priors that lead to projections of
outcomes for caribou that fall outside the observed range of variation
across the country, characterized here (Figure 2) by 95% predictive
intervals from a national analysis (Johnson et al. 2020), are
implausible. In other words, if we have seen no examples of
self-sustaining caribou populations at a particularly high level of
disturbance, it is unreasonable to project or expect this outcome. These
premises constrain the plausible demographic-disturbance priors, and
clarify that no additional information is needed to predict outcomes for
caribou where anthropogenic disturbance is high (Figure 5) and no
meaningful actions are taken to support recovery.

## Allowing for composition survey errors

Another novel aspect of this work is a term that allows for bias
associated with misclassification of adults and missing calves in
composition surveys [@ellington_divergent_2020;
@decesare_estimating_2012]. Lacking information about the true magnitude
of this bias, we examined sensitivity to a range of assumed values, and
found that adding unknown bias did increase uncertainty, but these
increases were not large (Supplement E). A key reason for this result is
that we assumed the same range of values for the probability of
misclassifying adults and the probability of missing calves. Presuming
that surveyed groups contain more cows than bulls, and that there is an
equal probability of misclassifying cows and young bulls and vice versa,
then misclassification errors cause high estimated recruitment. Missing
calves causes low estimated recruitment, and together these errors tend
to cancel one another out. We did not investigate sensitivity to the
assumption that new recruits are 50% females, but we note that this is
another possible source of error that would most likely lead to low
recruitment estimates.

Given lack of knowledge about true error rates, the most appropriate
interpretation of our results is that better information on composition
survey error rates is required in order to assess their importance.
Fecal pellets are proving to be an interesting source of information on
sex and age ratios in surveyed groups [@mcfarlane_genetic_2018;
@jones_novel_2023]; for example, in analysis of 14 populations of boreal
and southern mountain caribou, Jones et al. 2023 found that sampled
groups contained more females than males, and that mothers tend to stay
with their daughters and sons, while fathers do not. These results are
broadly consistent with our assumption that composition survey groups
include mothers and younger animals of both sexes that can be difficult
to classify visually. Documenting animals that are difficult to classify
as “unknown” can help analysts minimize and account for classification
bias [@ellington_divergent_2020], as can comparing classification
results from multiple independent observers. The probability of missing
calves is more difficult to quantify and account for using aerial survey
methods alone, but likely increases with vegetation cover.

Alternatives to aerial composition surveys for estimating recruitment
include fecal pellet analysis [@mcfarlane_genetic_2018;
@mcfarlane_spatial_2021; @mcfarlane_identification_2022] and calf
collaring [@ellington_divergent_2020]. Each method has costs and
benefits, and the reality is that monitoring everything everywhere all
the time in great detail is neither feasible nor necessary. We hope that
a bias correction term for the simple recruitment-mortality model
[@hatter_moose_1991; @decesare_estimating_2012;
@serrouya_comparing_2017; @eacker_web-based_2019; @eccc_scientific_2011;
@johnson_science_2020] that is used for many herd viability assessments
will enable and encourage practitioners to examine the sensitivity of
their conclusions to plausible composition survey error rates, and help
clarify thinking about where, when, and why additional information is
needed.

## Limitations and Future Possibilities

The main objective of Johnson et al's [-@johnson_science_2020;
@eccc_scientific_2011] national analysis was to inform a national
disturbance policy that applies to all ranges except SK1, and to inform
a different policy for SK1 [@eccc_scientific_2011; @eccc_boreal_2019]. A
hierarchical analysis [@leasure_hierarchical_2019] would be better
suited for estimating variation in demographic model parameters among
populations, and distinguishing that from variation among years.
However, data owners have not authorized the use of the data for that
purpose. To develop a proof-of-concept model with available published information we
constrained our example trajectory simulations by assuming that
populations remain in their quantile, so populations that are lower than
or higher than average remain lower or higher than average over time.
The Bayesian model is not similarly constrained, and can accommodate
evidence of a shift from unexpectedly low to unexpectedly high
demographic rates over time, or vice versa. Switching from low to high
each year is unlikely, but there is evidence of variation in the slope
of demographic-disturbance relationships among populations
[@rudolph_demographic_2017] so it is important that the Bayesian model
allows for a greater range of possibilities than our more constrained
simulated example trajectories. Although our simulated trajectories do
not include all plausible possibilities we believe they are sufficient
to demonstrate how a Bayesian model can be used to integrate local data
with national priors, and to show how monitoring requirements depend on
landscape condition.

For this analysis we used Kaplan-Meier estimates of observed adult female survival probability $\hat{S}_t$ and associated precisions $\tau_t$ [as in @eacker_web-based_2019; @lamb_effectiveness_2024], and assumed Beta distributed interannual variability [as in @johnson_science_2020], but we acknowledge there are other possible approaches [e.g. @dalgarno_bboutools_2024]. We believe the proof-of-concept model presented in this paper is adequate for making general points about the integration of national
demographic-disturbance relationships with local data, and accounting for
composition survey errors. However, in future we intend to use the more fully developed and documented 'bboutools' R package [@dalgarno_bboutools_2024] as a starting point for work with local decision makers to develop tools they can use inform local monitoring decisions.

Aspatial demographic models generally assume that observed animals are a
random and representative sample of a fully mixed population in which
survival and recruitment rates apply equally to all animals. However,
metrics derived from fecal pellet analysis such as reproductive success
[@mcfarlane_genetic_2018; @mcfarlane_spatial_2021;
@mcfarlane_identification_2022], age structure and pregnancy rates
[@flasko_fecal_2017; @mcfarlane_identification_2022], and
mother-offspring co-occurence [@jones_novel_2023] suggest high variation
in fitness among individuals [@mcfarlane_genetic_2018] and spatial
structure within populations [@jones_novel_2023;
@mcfarlane_spatial_2021; @mcfarlane_identification_2022]. Analyses of
movement data also show spatial variation in risks and benefits for
caribou and variation in habitat preferences and space use among
individuals [e.g. @decesare_linking_2014; @avgar_spaceuse_2015].
Although there is evidence of spatial structure and variation in
demographic rates among individuals within populations, for most caribou
populations there is not enough data available to estimate spatial
variation in demographic model parameters.

A number of additional complexities can be incorporated into aspatial
Bayesian population models, including variation in demographic rates
among individuals [@armstrong_incorporating_2021], population size
information from aerial or fecal dna surveys, and additional age classes
[@moeller_integrating_2021; @mcnay_demographic_2022]. However, in many
contexts, population size and fecal dna survey information is not
available, and a framework that includes these sources of information
will be unnecessarily complicated. At present, our model is intended as
a proof-of-concept to show the value of integrating national
demographic-disturbance relationships with local data, accounting for
composition survey errors, and assessing the value of local monitoring
information. Rather than further complicate a proof-of-concept model
applied to simulated data, our next step will be to collaborate with an
interested decision maker to develop a tool informed by this analysis
that meets their real local needs, recognizing that these needs will
vary among monitoring programs [@nbckc_nbckc_2021; @nbckc_nbckc_2021].

## Assessing the value of additional monitoring

We applied the framework proposed by Dunham et al. (2023) to show that
no additional local monitoring information is needed to predict whether
boreal caribou populations in highly disturbed landscapes are
self-sustaining. There is more variation in caribou demographic rates
among study areas with low to moderate disturbance, and thus more need
for local data to inform demographic projections. In these places, our
analysis suggests that there may be diminishing returns from
increasing monitoring effort, but more information on monitoring costs,
objectives, and existing local data would be required to assess the
adequacy and cost-effectiveness of monitoring scenarios in particular
circumstances. Monitoring costs vary substantially with technologies
used, the remoteness of herds, and the difficulty of collaring animals,
which in turn depends on the number of currently collared animals, the
population size, and vegetation across the landscape
[@nbckc_boreal_2021]. It is important to recognize that projecting
population viability is only one possible objective for monitoring
[@nbckc_boreal_2019], and that the impacts of monitoring on threatened
animals should be considered [@nbckc_boreal_2019]. Since monitoring
costs, options and goals are context specific, we believe the most
effective way forward will be to use this proof-of-concept model and
analysis as a starting point for working with local decision makers to
build decision support tools suited to their particular circumstances,
needs and questions.

# Acknowledgements

# Author contributions
**Josie Hughes:** Conceptualization, Formal analysis, Methodology, Investigation,
Supervision, Software, Writing – original draft, Writing – review and editing, 
Visualization
**Sarah Endicott:** Resources, Software, Writing – review and editing
**Anna M Calvert:** Conceptualization, Writing – original draft, Writing – review and editing
**Cheryl A Johnson:** Conceptualization, Writing – review and editing

<!-- The following line inserts a page break  -->

\newpage

# References

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

::: {#refs}
:::

### Colophon

This report was generated on `r Sys.time()` using the following
computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
#if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
#if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```