### 2025-02-21 analysis/paper/paper.Rmd
preparing EI submission

---
title: "Integration of national demographic-disturbance relationships and local data can improve caribou population viability projections and inform monitoring decisions"
author:
  - name: Josie Hughes
    email: josie.hughes@ec.gc.ca
    institute: [ECCC]
    correspondence: true
  - name: Sarah Endicott
    email: sarah.endicott@ec.gc.ca
    institute: [ECCC]
    correspondence: false
  - name: Anna M Calvert
    email: anna.calvert@ec.gc.ca
    institute: [ECCC]
    correspondence: false
  - name: Cheryl A Johnson
    email: cheryl-ann.johnson@ec.gc.ca
    institute: [ECCC]
    correspondence: false

institute:
  - ECCC: Landscape Science and Technology Division, National Wildlife Research Center, Environment and Climate Change Canada, 1125 Colonel By Dr, Ottawa, ON, K1S 5B6, Canada
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      number_sections: no
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography: CaribouDemographicModel.bib
csl: "../templates/ecological-informatics.csl" # Insert path for the bib-style
abstract: |
  Across fifty-eight boreal caribou study areas in Canada, both survival and recruitment decrease with the percentage of the study area that is disturbed. There is variation in demographic rates among study areas, particularly where anthropogenic disturbance is low, but no populations inhabiting areas with high anthropogenic disturbance are considered self-sustaining. Demographic projections derived from local population-specific data are uncertain for populations with limited monitoring. We propose a simple Bayesian population model that integrates prior information from a national analysis of demographic-disturbance relationships with available local demographic data to improve population viability projections, and to reduce the risk that a lack of local data will be used as a reason to delay conservation action. The model also acknowledges additional uncertainty and potential bias due to misidentification of sex or missing calves, through a term derived from a simple model of the recruitment survey observation process. We combine this Bayesian model with simulations of plausible population trajectories in a value of information analysis framework to show how the need for local monitoring varies with landscape condition, and to assess the ability of alternative monitoring scenarios to reduce the risk of errors in population viability projections. Where anthropogenic disturbance is high, reasonably accurate status projections can be made using only national demographic-disturbance relationships. At lower disturbance levels where initial uncertainty is high, each additional year of monitoring provides less new information. The estimated probability of viability indicates whether more information is needed to improve accuracy of population viability projections. 
keywords: |
  woodland caribou (*Rangifer tarandus caribou*); anthropogenic disturbance scenarios; population viability; value of information; monitoring decisions; species at risk; prior information
highlights: |

  * Across fifty-eight boreal caribou study areas in Canada, both survival and recruitment decrease with the percentage of the study area that is disturbed.

  * We show how Bayesian integration of local data and prior information from a national analysis of demographic-disturbance relationships can reduce uncertainty about the viability of sparsely monitored populations.

  * Integrating prior information about demographic-disturbance relationships clarifies that no additional information is needed to project population viability where disturbance is high and there are no actions to support recovery.

  * At lower disturbance levels where uncertainty is higher there are diminishing returns on additional monitoring, and the estimated probability of viability indicates whether more information is needed.

  * We introduce a method for acknowledging potential errors in recruitment surveys to help clarify where and when these errors are likely to alter conclusions about population viability.

editor_options: 
  chunk_output_type: inline
  markdown: 
    wrap: 72
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

<!-- With the following code you can access and display values from the yml header above. -->

<!-- Link to data here: https://drive.google.com/file/d/179lw6MPK4l3xGmded9kofzJkB6VlJVb9/view?usp=sharing -->

Keywords: `r rmarkdown::metadata$keywords`

Highlights: `r rmarkdown::metadata$highlights`

<!-- The following code chunk defines some general settings how code chunks should behave. -->

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figs/",
  dpi = 300
)
```

<!-- The actual document text starts here: -->

# Introduction

The number of species threatened with extinction is high and getting higher [@ipbes_global_2019]. Reliable information on the status of populations and likely impacts of management are needed to inform conservation action, but effective monitoring is challenging and costly [@legg_why_2006; @lindenmayer_effective_2018; @wintle_allocating_2010;@dunham_strategic_2023]. Resources to support conservation and recovery are limited [@buxton_balancing_2022; @mccarthy_financial_2012;@gerber_conservation_2016; @wintle_spending_2019], delays in action increase risk [@martin_acting_2012], and some populations have been monitored to extinction [@lindenmayer_counting_2013]. A recent analysis of action plans for Canadian species at risk found that half of tasks are research and monitoring, and this proportion is highest for species with higher risk of extinction, highlighting a need to more carefully consider when, where, and why additional information is needed to inform conservation action [@buxton_balancing_2022]. Analysis of the implications of monitoring outcomes can help to clarify whether additional monitoring is necessary to inform conservation action [@bennett_when_2018; @dunham_strategic_2023; @runge_which_2011;@gregory_structured_2012].

Boreal caribou are widely and sparsely distributed through the Canadian boreal region across a gradient of anthropogenic disturbance [@eccc_scientific_2011; @johnson_science_2020] that includes some of the most intact and inaccessible places remaining on earth [@hirsh-pearson_canadas_2022; @ibisch_global_2016;
@venter_sixteen_2016]. Across 58 boreal caribou study areas, both survival and recruitment decrease with the percentage of the area that is disturbed, and no populations inhabiting areas with high anthropogenic disturbance are considered viable in the absence of interventions such as predator control or maternal penning (i.e. self-sustaining) [@eccc_scientific_2011; @johnson_science_2020]. Identifying populations that are declining is important for clarifying where and when conservation action is needed. Disturbance affects caribou demography both through direct impacts on habitat availability and use, and indirect impacts on predators and alternate prey [@nbckc_boreal_2022; @eccc_scientific_2011]. Decreasing recruitment with increased disturbance has also been observed over time where historic demographic data exist [@rudolph_demographic_2017; @hervieux_widespread_2013], but many places lack sufficient data to characterize local demographic-disturbance relationships [@johnson_science_2020]. There is also variation in demographic rates among study areas, particularly where anthropogenic disturbance is low [@johnson_science_2020; @neufeld_primary_2021;
@fortin_forest_2017; @gagne_impact_2016], which leads to uncertainty in projections that rely only on national-scale relationships [@stewart_climate-informed_2023; @dyson_existing_2022].

Analysis methods that can make better use of available monitoring data to inform conservation decisions include Bayesian population modelling [@schaub_integrated_2021] and value of information analyses [@bennett_when_2018; @dunham_strategic_2023; @runge_which_2011;@gregory_structured_2012]. For sparsely monitored populations, incorporating prior information from elsewhere can help reduce uncertainty [@drummond_combining_2018; @parlato_incorporating_2021; @tufto_bayesian_2000]. Various efforts have been made to improve caribou population viability projections and growth rate estimates using Bayesian methods [@moeller_integrating_2021; @mcnay_demographic_2022;@eacker_web-based_2019; @lamb_effectiveness_2024; @dalgarno_bboutools_2024]. An advantage of Bayesian demographic models is that they account for multiple sources of uncertainty [@schaub_integrated_2011; @eacker_assessing_2017;@schaub_integrated_2021]. However, Bayesian caribou population models developed to-date [@moeller_integrating_2021; @mcnay_demographic_2022;@eacker_web-based_2019; @lamb_effectiveness_2024; @dalgarno_bboutools_2024] have not included prior information about the observed distribution of demographic rates across the country that could reduce uncertainty about the status of poorly monitored populations.

Here, we offer a simple Bayesian population model for boreal caribou (*Rangifer tarandus
caribou*) that integrates prior information about the impacts of landscape change [i.e. demographic-disturbance relationships in @johnson_science_2020] with available demographic data to reduce uncertainty in population viability projections. Note that for the purposes of this analysis we consider a population to be viable if the population growth rate is > 0.99. We also offer a novel method of acknowledging and accounting for errors in composition surveys. Following a value information analysis framework for conservation status assessments [@dunham_strategic_2023], we combine this Bayesian model with simulations of plausible population trajectories [@stewart_climate-informed_2023; @dyson_existing_2022] to (i) show how the value of local monitoring varies with landscape condition, and (ii) assess the ability of alternative monitoring scenarios to reduce the risk of errors in population viability projections.

Our proposed model represents an important improvement over less formally integrated methods for estimating and projecting impacts of disturbance on population viability [@eccc_scientific_2011; @johnson_science_2020] because it accounts for uncertainties in both local data and demographic-disturbance relationships. We show that integrating the most informative predictor of outcomes for boreal caribou across their range can substantially reduce uncertainty in local demographic projections and improve the accuracy of population viability projections on changing landscapes. This approach can help clarify where available information is – and is not – sufficient for assessing and projecting population viability, and thereby reduce the risk that uncertainty is used as a reason to delay conservation action for imperiled populations [@bissonnette_controverses_2020; @boan_climate_2018]. Our approach is tailored to boreal caribou, but could be adapted for other species at risk found in multiple populations that are costly to monitor and sensitive to habitat disturbance.

# Methods

## Overview of methods and notation

To investigate how the accuracy and precision of predictions vary with both landscape condition and monitoring effort, we analyze simulated data using a Bayesian model with informative priors (section 2.4) and a composition survey bias term (section 2.1). Steps in our analysis procedure include defining anthopogenic disturbance and monitoring scenarios (section 2.2), simulating population dynamics and monitoring (section 2.3) to produce example demographic trajectories for each scenario, and analyzing example demographic trajectories with a Bayesian model that enables integration of (simulated) local demographic data with prior knowledge of demographic-disturbance relationships (section 2.4). Outcomes of interest include expected population growth rate, the probability of correctly assessing population status, the posterior probability of population viability, and the expected value of local information (section 2.5). We refer to the model used to simulate example trajectories as the "national" model because parameters are from a national analysis of demographic-disturbance relationships [@johnson_science_2020]. We distinguish this national model from the Bayesian model that is used to analyze simulated example demographic trajectories. Where appropriate the national model and the Bayesian model share the same structure, and we have selected notation to highlight the similarities. Bayesian model estimates are distinguished from observed values, estimated expected values, true values, and true expected values using $\hat{\text{ }}$ $\tilde{\text{ }}$ $\dot{\text{ }}$ and $\bar{\text{ }}$ symbols respectively; for example, $S_t$ is the Bayesian estimate of survival in year $t$, $\hat{S}_t$ is observed survival, $\tilde{S}_t$ is the Bayesian estimate of expected survival (without stochastic interannual variation), $\dot{S}_t$ is true survival, and $\bar{S_t}$ is true expected survival (without stochastic interannual variation). For reference, all variables and parameters are defined in supplemental Table A1.

```{r setup2}
library(ggplot2)
library(dplyr)
library(stringr)
library(tidyr)
library(purrr)
library(ggpubr)
library(MetBrewer)
library(here)
library(tmap)
library(sf)

devtools::load_all(here::here())

#set paths
data_path_raw <- "analysis/data/raw_data/"
data_path_drvd <- "analysis/data/derived_data"

pal_nm <- "Redon"

fig_widths <- c(min = 1.18, single = 3.543, mid = 5.51, two = 7.48)

runScenarios = F #To do: set up to run full analysis from within this document.

resultScenario = "S13"
resultSet ="cmult6ay1aSf1"
figSet = c("diffslow","distScns","EVSI","power","EVPIall","powerLam","probViability")

erSums = read.csv(here::here(paste0("tabs/",resultScenario,"/ErrorsByBandcmult6ay1aSf1.csv")))

if(runScenarios){
    source(here::here("analysis/scripts/0_nationalModelTrajectories.R"))
    source(here::here("analysis/scripts/0_bayesianModelExamples.R"))
    #create scenarios with sensitivity1.R and run in cloud
    #run viewSensitivityResults.R for S8 and S7
    #run EVSIPlots.R
}

for(f in figSet){
  if(f =="EVPIall"){addBit=""}else{addBit=resultSet}
  currentFile = here::here(paste0("figs/",resultScenario,"/",f,addBit,".png"))
  newFile = here::here(paste0("figs/forPaper/",f,addBit,".png"))
  file.copy(currentFile, newFile,overwrite=T)

  currentFile = here::here(paste0("figs/",resultScenario,"/",f,addBit,".pdf"))
  newFile = here::here(paste0("analysis/paper/figs_submit/",f,addBit,".pdf"))
  file.copy(currentFile, newFile,overwrite=T)
}
```

## Caribou population monitoring data and composition survey errors

The most widely available source of boreal caribou demographic
information is from cows fitted with telemetry collars [e.g.
@courtois_aerial_2003; @rettie_dynamics_1998; @hervieux_widespread_2013;
@decesare_estimating_2012], though genetic methods are playing an
increasingly important role in some caribou monitoring programs [e.g. @flasko_fecal_2017;
@hettinga_estimating_2012; @mcfarlane_genetic_2018;
@moeller_integrating_2021; @jones_novel_2023;
@mcfarlane_identification_2022; @mcfarlane_spatial_2021]. Caribou cows
are typically captured via net-gun from aircraft in late winter, and
affixed with collars that send a mortality signal when no longer
active [see @nbckc_nbckc_2021], thus allowing estimation of adult
survival [e.g. @dalgarno_bboutools_2024; @mcloughlin_declines_2003;
@hervieux_widespread_2013; @eacker_web-based_2019; @lamb_effectiveness_2024]. Recruitment is estimated from
aerial surveys of the composition of groups of animals that include
collared animals [@nbckc_nbckc_2022]. Groups typically also contain uncollared animals, and a critical assumption in these surveys is that different demographic groups are equally detectable
[@ellington_divergent_2020]. Errors that can bias composition survey
results include misidentifying young bulls as cows or vice versa , and
failing to detect calves [@ellington_divergent_2020;
@decesare_estimating_2012]. Survival and recruitment survey data can be combined in a simple
recruitment-mortality population model [@hatter_moose_1991;@hatter_revisiting_2020] that serves
as the basis for many herd viability assessments
[@decesare_estimating_2012; @serrouya_comparing_2017;
@eacker_web-based_2019; @eccc_scientific_2011; @johnson_science_2020; @dalgarno_bboutools_2024].

While we believe that using available information to reduce uncertainty
is important, we also believe it is important to acknowledge remaining
uncertainty. In general, parametric Bayesian population models include a
number of simplifying assumptions, and misspecified models can yield
biased or overconfident results [@schaub_integrated_2021]. More
specifically, models that do not account for observation errors can be
misleading [@schaub_integrated_2021]. Although several authors have
pointed out that misclassification of adults and missing calves in
composition surveys can bias recruitment estimates from those surveys
[@ellington_divergent_2020; @decesare_estimating_2012], none have thus
far integrated these sources of error into Bayesian caribou population models [@moeller_integrating_2021; @mcnay_demographic_2022; @eacker_web-based_2019; @lamb_effectiveness_2024; @dalgarno_bboutools_2024]. We address this gap with a bias term derived from a simple model of the recruitment survey observation process.

To derive a composition survey bias term we assume each group
of animals in a calf:cow composition survey contains one or more
collared adult females ($T$), and may also include: uncollared adult
females misidentified as young bulls or unknown sex ($U$); correctly
identified uncollared adult females ($V$); young bulls correctly
identified as male or unknown sex ($O$); young bulls misidentified as
uncollared adult females ($P$); observed calves ($J$); and unobserved
calves ($K$). The apparent number of adult females in the group is
$T+V+P=Tw$, where $w$ is the observed ratio of the apparent number of
adult females to collared animals in the recruitment survey.
The ratio of young bulls to uncollared adult females in the group is
$q = (P+O)/(U+V).$ Assuming an equal probability $u$ of misidentifying
young bulls as adult females and vice versa, we get $V=(U+V)(1-u)$ and
$P=(O+P)u$. Given a probability $z$ of missing calves, we get
$J=(J+K)(1-z)$. Given these assumptions and definitions, the sex and
bias-adjusted recruitment rate $X$ can be written as a function of the
observed calf:cow ratio $R$, the cow multiplier $w$, the ratio of young
bulls to uncollared adult females $q$, and the misidentification probabilities $u$
and $z$ (see Supplement B for details). For simplicity, we combine the
probabilities into a bias term $c$ that is integrated into the national simulation model (section 2.4) and
the Bayesian model (section 2.5):

\begin{equation}
  c=\frac{w(1+qu-u)}{(w+qu-u)(1-z)}; X=cR/2.
  (\#eq:c)
\end{equation}

## Anthropogenic disturbance and monitoring scenarios

In simulated monitoring scenarios, we assume collars are deployed in January, and
continue to function for up to 6 years. The target number of collared females $n$ varies
among scenarios, as does the total duration of the monitoring program
$d$ and the ratio of observed adult females to
collared animals in the recruitment
survey $w$ (Table \@ref(tab:monitoringScenarios)). Collars lost to
mortality may be replaced each year ($o=1$), or there may be a 3 year
gap between deployments ($o=4$).

We focus on exploring a set of scenarios in which buffered anthropogenic
disturbance, measured as the percentage of area within 500 m of some
type of anthropogenic disturbance [@johnson_science_2020;
@eccc_scientific_2011], increases by 1% per year (Figure
\@ref(fig:figDisturbance)). The duration of monitoring prior to the
projection period varies among scenarios from 1 to 24 years. The
projection period begins in 2024, and we consider how uncertainty
changes as the projection period increases from 5 to 20 years. The
amount of anthropogenic disturbance at the beginning of the projection
period varies from low (0%) through medium-low (20%) and medium-high
(40%) to high (60%). These simple hypothetical scenarios span a large
range to show how the need for local monitoring to reduce uncertainty
depends on landscape state, not to represent the reality of landscape
change. Scenarios in which anthropogenic disturbance remains the same or
decreases over time are included in Supplement D.

```{r monitoringScenarios, echo=FALSE,message=FALSE}

source(here::here("analysis/scripts/formatScenarios.R"))

tbl <- tbl %>% separate_wider_delim(name, names = c("symbol", "name"), delim = ", ") %>% 
  select(-name, symbol, units, description, values)

knitr::kable(tbl,caption="Monitoring scenario parameters and values used in simulations.",
             col.names = str_to_title(colnames(tbl)))
```

```{=html}
<!--
*Monitoring WG input from BC, ON, NL, (MB pending): lots of detail provided but generally those numbers suggest we could try values of ~3-10 [e.g. 3,6,9?] for the cow multiplier, 5-50 for collars [e.g. 5,10,20,40,60?], [not sure about renewal interval, but see SaraM email for some insights]. We know (e.g. from the MonitoringWG perspectives report/update to Rettie summary) that number of years of monitoring can be as low as 2 or as many as 20+

*from Anna: withouot access to actual estimates of R and S for different ranges, some potentially interesting other ranges to compare to could be an Alberta range or two, where there are ++years of R and S estimates but really low lambda (E.g. ESAR and WSAR have 27+ yrs consecutive monitoring but declining numbers and high disturbance), some sections of NWT1 that are thought to be self-sustaining and have pretty good data sets (e.g. Dehcho North and South, Hay River Lowlands with 13+yrs of S and R) -could also use other MB ranges as examples for fairly sparse data (though not as bad as ON), and QC or LAB ranges as examples of good data (e.g. how about Charlevoix for a very well monitored but tiny pop, northern QC for really broad-expanse pop, or Labrador pops have long term monitoring data but uncertain status...)

*feedback from MonitoringWG contacts includes details on collar numbers for BC, ON, NL, (MB pending); e.g. in Ontario they aim for 20-30 collars in smaller pops, 35+ in bigger pops; Labrador aims to have between 5-25 collars monitored at any one time, depending on pop size; in BC, real numbers of collared animals have ranged been between 3-59 per pop/year, varying both across-yrs-within-pop and across-pop-within-yr [with COVID yrs by far the lowest].

relevant refs: [@rettie_summary_2017];?

TO DO: Fire scenarios? Add high fire scenario in appendix of one by one sensitivity checks.

TO DO: run low initial population size scenario, put in appendix of one by one sensitivity checks? 
-->
```
## Simulation of population dynamics and monitoring

To investigate the effectiveness of various monitoring strategies, and to show how this
varies with landscape condition, we begin by simulating example demographic trajectories and observations of those trajectories using a modified version of Johnson et al's [-@johnson_science_2020] demographic model described by Dyson et al. [-@dyson_existing_2022]. For brevity, we refer to this as the "national" model, and for clarity we include a full description in the following paragraphs. Together, a set of example trajectories from the national model represent alternative hypotheses about the true state of a particular caribou population. To assess the accuracy of population status assessments and the expected value of alternative monitoring strategies [@runge_which_2011; @dunham_strategic_2023] we analyze each trajectory of simulated observations using the Bayesian model (section 2.4). We also use the sets of simulated trajectories to confirm that prior predictions from our Bayesian model are consistent with knowledge of the observed distribution of outcomes for caribou across the country [@johnson_science_2020].

Given a population of post-juvenile females at the beginning of year $t$,
$\dot{N}_t$, the number that survive to the census, $\dot{W}_t$, is
binomially distributed with true survival probability $\dot{S}_t$:
$\dot{W}_{t} \sim \text{Binomial}(\dot{N}_t,\dot{S}_t)$. Realized recruitment rate varies with population density [@lacy_vortex_2017], and the number of juveniles recruiting to the
post-juvenile class at the census $(\dot{J}_{t})$ is a binomially
distributed function of the number of surviving post-juvenile females ($\dot{W}_t$)
and the maximum potential recruitment of female calves per cow ($\dot{X}_t$):

\begin{equation}
  \dot{J}_{t} \sim \text{Binomial}(\dot{W}_t,\dot{X}_t[p_0-(p_0-p_k)(\frac{\dot{W}_t}{k})^b]\frac{\dot{W}_t}{\dot{W}_t+a}),
  (\#eq:dotJ)
\end{equation}

Given the values of parameters $p_0,p_k,a,$ and $b$ used by Johnson et
al. (2020) (Table A1), recruitment rate is lowest $(0.5\dot{X}_t)$ when
$\dot{N}_t=1$, approaches a maximum of $\dot{X}_t$ at intermediate
population sizes, and declines to $0.6\dot{X}_t$ as the population
reaches carrying capacity of $k=10000$.
The post-juvenile female population in the next year includes both
survivors and new recruits: $\dot{N}_{t+1}=\text{min}(\dot{W}_t+\dot{J}_t,r_{max}\dot{N}_t)$, where the maximum potential population growth rate is $r_{max}=1.3$.

The observable calf:cow
ratio $\dot{R}_t$ is adjusted for sex ratio and composition survey errors
(Supplement B) to get the maximum
potential recruitment of female calves per cow $\dot{X}_t$:

\begin{equation}
  \dot{X}_t=\dot{c}\dot{R}_t/2;\dot{c}=\frac{w(1+\dot{q}\dot{u}-\dot{u})}{(w+\dot{q}\dot{u}-\dot{u})(1-\dot{z})},
  (\#eq:dotX)
\end{equation}

where $w$ is the ratio of observed adult females to
collared animals, $\dot{q}$ is the ratio of young bulls to adult
females, $\dot{u}$ is the probability of misidentifying young bulls as
adult females and vice versa, and $\dot{z}$ is the probability of
missing a calf. We examine sensitivity to these unknown parameters by
selecting values for each simulated example population from plausible
ranges: $\dot{u}$ and $\dot{v}$ vary uniformly between 0 and 0.2, and
$\dot{q}$ varies uniformly between 0 and 0.6. The distinction between the true composition bias term $\dot{c}$ and the term $c$ we include in the Bayesian model (eq \@ref(eq:c)) is that the (simulated) truth is known, and the Bayesian model bias is uncertain.

Interannual variation in survival and recruitment is modelled using
truncated beta distributions [rtrunc function;
@novomestky_package_2016]:
$\dot{R}_t \sim \text{TruncatedBeta}(\bar{R}_t,v_R,l_R,h_R); \dot{S}_t \sim \text{TruncatedBeta}(\bar{S}_t,v_S,l_S,h_S)$.
Coefficients of variation among years $(v_R,v_S)$ and
maximum/minimum values $l_R,h_R,l_S,h_S$ for recruitment and survival
are given in Table A1. Expected recruitment ($\bar{R}_t$) and survival
($\bar{S}_t$) vary with disturbance according the beta regression models
estimated by Johnson et al. [-@johnson_science_2020]:

\begin{equation}
\bar{R}_t \sim Beta(\mu^R_t,\phi^R); 
log(\mu^R_t) = \dot{\beta}^R_0+\dot{\beta}^R_a A_t+\dot{\beta}^R_f F_t,
  (\#eq:barR)
\end{equation}

\begin{equation}
\bar{S}_t \sim (46 \times Beta(\mu^S_t,\phi^S)-0.5)/45;
log(\mu^S_t) = \dot{\beta}^S_0+\dot{\beta}^S_a A_t.
  (\#eq:barS)
\end{equation}

$\phi^R \sim \text{Normal}(19.862,2.229)$ and
$\phi^S \sim \text{Normal}(63.733,8.311)$ are precisions of the Beta
distributed errors [@ferrari_beta_2004]. At the beginning of a
simulation for an example population, regression coefficient values are
sampled from Gaussian distributions (see Table \@ref(tab:priors)) and
the population is assigned to quantiles of the Beta error distributions
for survival and recruitment. The population remains in these quantiles
as disturbance changes over time, so there is substantial persistent
variation in recruitment and survival among example populations (Figure
\@ref(fig:exampleTrajectories)).

In our simulated example populations the number of collared cows at the time of the composition survey $\hat{T}_t$ depends on the true survival rate $\dot{S}_t$ and the number of collared cows at the beginning of the year $\hat{I}_t$: $\hat{T}_t \sim \text{Binomial}(\hat{I}_t,\dot{S}_t)$. The number of cows in the composition survey $\hat{W}_t$ is given by the number of collared cows $\hat{T}_t$ and the apparent number of adult
females per collared female observed in the composition survey $w$: $\hat{W}_t = w\hat{T}_t$. The number of observed calves $\hat{J}_t$ also depends on the unadjusted apparent recruitment for the population $\dot{R}_t$:

\begin{equation}
\hat{J}_t \sim \text{Binomial}(\hat{W}_t,\dot{R}_t).
  (\#eq:hatJ)
\end{equation}

To ensure that results are comparable across monitoring scenarios and that populations remain extant through the
monitoring period, all simulations are started with $N_0=5000$ animals
24 years prior to the start of the projection period. Realized population growth rate is $\dot{\lambda}_t=\dot{N}_t/\dot{N}_{t-1}$ Expected population growth rate (without interannual variation, demographic stochasticity, or effects of population density) is $\bar{\lambda}_t=\bar{S}_t(1+\dot{c}\bar{R}_t/2)$.

## Bayesian integration of local demographic data and national disturbance-demographic relationships

Local demographic and covariate data provided as inputs to the Bayesian
population model include the number of collared cows at the start of the year ($\hat{I}_{t}$), the number of collared animals that survive the year ($\hat{T}_{t}$), the observed number
of adult females and calves in the recruitment surveys
($\hat{W}_t, \hat{J}_t$), the apparent number of adult females per
collared female in the composition survey $w$, and the measures of
anthropogenic disturbance and fire ($A_t, F_t$). Recruitment rate $R_t$
is estimated from the observed number of calves $\hat{J}_t$ and adult
female caribou $\hat{W}_t$ assuming a Binomial distribution:
\begin{equation}
\hat{J}_t \sim \text{Binomial}(\hat{W}_t,R_t).
  (\#eq:hatJEst)
\end{equation}

The difference between equations \@ref(eq:hatJ) and \@ref(eq:hatJEst) is that in the simulation model an
unadjusted apparent recruitment rate $\dot{R}_t$ is selected and used to
simulate observations. In the Bayesian model, in contrast, observations
are provided, and used to estimate the unobserved recruitment rate
$R_t$.

In both models recruitment probability is a function of anthropogenic
disturbance and fire with a log link and Beta distributed interannual variation [@johnson_science_2020;
@stewart_climate-informed_2023; @dyson_existing_2022]. The coefficient of variation $\nu_R$ is constant, while the mean $\tilde{R_t}$ and standard deviation $\sigma_{R,t}$ vary with disturbance:
\begin{equation}
log(\tilde{R_t})=\beta^R_0+\beta^R_a A_t+\beta^R_f F_t
  (\#eq:Rtilde)
\end{equation}

\begin{equation}
R_t \sim \text{Beta}(\tilde{R_t},\sigma_{R,t});
\sigma_{R,t} = min(\nu_R \tilde{R_t},\sqrt{\tilde{R_t} (1-\tilde{R_t})})
(\#eq:R)
\end{equation}

In the simulation model (eq \@ref(eq:barR)), regression coefficient
values ($\dot{\beta}^R_0,\dot{\beta}^R_a,\dot{\beta}^R_f$) are selected
for each example population and used to calculate recruitment rate. In
the Bayesian model (eq \@ref(eq:R)), unobserved regression coefficients
($\beta^R_0,\beta^R_a,\beta^R_f$) are estimated from local recruitment
survey and disturbance observations.

To account for sex ratio of calves and additional uncertainty caused by composition survey errors (Supplement B) estimated recruitment is adjusted to get expected female
recruits per female $X_t=cR_t/2$. The composition survey bias term $c$ (eq \@ref(eq:c)) depends on the known
apparent number of adult females per collared animal $w$, the unknown
ratio of young bulls to adult females $q$, the unknown probability of
misidentifying young bulls as adult females and vice versa $u$, and the
unknown probability of missing a calf $z$ (Supplement B). Prior
uncertainty about the value of the bias term $c$ is
approximated with a Log-normal distribution (Supplement B). Given the
apparent number of adult females per collared animal $w$ in each
monitoring scenario (Table \@ref(tab:monitoringScenarios)), we calculate
the mean and standard deviation of 10,000 samples of $\log{c}$ assuming
the ratio of young bulls to adult females $q$ varies uniformly between 0
and 0.6, the adult misidentification probability $u$ varies uniformly
between 0 and 0.2, and the probability of missing a calf $z$ varies
uniformly between 0 and 0.2.

Survival $S_t$ is estimated from the observed number of collared cows at the beginning of each year $\hat{I}_t$ and the number of those that survive the year $\hat{T}_t$ assuming a Binomial distribution: $\hat{T}_t \sim \text{Binomial}(\hat{I}_t,S_t)$. Note that real known-fate radio collar data should be analyzed with a method that accounts for variation in survival rate and the number of collared animals throughout the year [e.g. @dalgarno_bboutools_2024; @eacker_web-based_2019; @lamb_effectiveness_2024]. However, our implementation of a two-step non-parametric survival analysis method [following @eacker_web-based_2019] yielded slightly biased survival estimates.

As done by Johnson et al. [-@johnson_science_2020], we model survival
probability as an adjusted function of buffered anthropogenic
disturbance $A_t$ with a log link and Beta distributed interannual variation with coefficient of variation $\nu_S$, mean $\tilde{S_t}$ and standard deviation $\sigma_{S,t}$ [@stewart_climate-informed_2023;
@dyson_existing_2022]: 
\begin{equation}
\tilde{S_t}=(46 e^{\beta^S_0+\beta^S_a A_t}-0.5)/45
  (\#eq:Stilde)
\end{equation}

\begin{equation}
S_t \sim \text{Beta}(\tilde{S_t},\sigma_{S,t});
\sigma_{S,t} = min(\nu_S \tilde{S_t},\sqrt{\tilde{S_t} (1-\tilde{S_t})})
(\#eq:S)
\end{equation}

Again assuming a simple demographic model with a
census in late winter of surviving females and new recruits to the
post-juvenile class [@hatter_moose_1991; @hatter_revisiting_2020], the expected population growth rate (without interannual variation or demographic stochasticity) is $\tilde{\lambda}_t=\tilde{S}_t(1+c\tilde{R}_t/2)$.

All regression coefficients are assumed to be Gaussian distributed, with
prior means from Johnson et al's (2020) regression models
(Table \@ref(tab:priors)). We calibrated the standard deviations of the
priors so that the the 95% prior predictive intervals for survival and
recruitment from the Bayesian model match the 95% prediction intervals
from Johnson et al's regression model (Supplement C, Figure
\@ref(fig:examplePredictions)). To show the value of informative priors, we also include a weakly informative scenario with prior means of the intercept parameters from Dalgarno et al. [-@dalgarno_bboutools_2024] ($\beta^R_0=-0.3689, \beta^S_0=-0.2126776$), prior means of the slope parameters set to zero, and standard deviations for all parameters (Table \@ref(tab:priors)) multiplied by ten.

```{r priors, echo=FALSE,message=FALSE}

source(here::here("analysis/scripts/formatPriors.R"))

knitr::kable(tbl,caption="Prior means and standard deviations of Gaussian distributed survival and reproduction model parameters. Standard deviations are calibrated so that the 95\\% prior predictive intervals for survival and recruitment from the Bayesian model match the 95\\% prediction intervals from Johnson et al's regression model (Supplement C).")
```

For reference, all variables and parameters are defined in supplemental Table A1. Models and analysis were implemented using the R language
[@r_core_team_r_2023] and a variety of *tidyverse* packages
[@wickham_welcome_2019]. We use the open source Gibbs sampling software
JAGS [@plummer_jags_2023] and the *R2jags* R package [@su_r2jags_2021]
to fit the Bayesian model. We ran 15,000 iterations of 4 chains,
discarded the first 10,000 interations, and used a thinning rate of 2.
We assessed convergence for a small random subset of example
trajectories using a threshold for the Gelman-Rubin diagnostic
[@gelman_inference_1992] of r-hat \< 1.1, and visually inspected
posterior autocorrelation and traceplots produced by the *mcmcplots* R
package [@curtis_mcmcplots_2018]. Code required to run analyses and
reproduce figures in this paper is available at
<https://github.com/LandSciTech/Caribou-Demographic-Projection-Paper>,
and in the *caribouMetrics* R package available on GitHub at
<https://github.com/LandSciTech/caribouMetrics/tree/EI-paper-submission>.

## Probability of correct status assessment and expected value of information

In order to assess whether a monitoring program is likely sufficient to
inform decisions, it is important to select outcome measures that are
relevant to decision making [@gregory_structured_2012]. For boreal
caribou and other endangered species, a relevant outcome is the
probability of correctly predicting the true population status
(increasing or decreasing) [@dunham_strategic_2023]. To calculate this
outcome for each disturbance and monitoring scenario (Fig
\@ref(fig:figDisturbance)) we simulate observations from 500 example trajectories (e.g. Fig
\@ref(fig:exampleTrajectories)), fit the Bayesian model to each of these, and calculated the percentage of these
500 examples in which the Bayesian posterior expected population growth rate
and the true expected population growth rate are both greater than 0.99, or both less than or equal to 0.99.

Following Dunham et al. [-@dunham_strategic_2023] we also
calculate the expected value of perfect information (EVPI) and the
expected value of sample information (EVSI) [@runge_which_2011;
@yokota_value_2004]. We assign a value or utility of
$\eta_{\alpha,\psi}=1$ when the true state of the population $\psi$
matches the assigned status $\alpha$, and value of $\eta_{\alpha,\psi}=0$
when the status assessment is incorrect. This is "the value ascribed by
the decision-maker to the outcome, and is thus the measure of management
performance" [@runge_which_2011]. The expected value of current
information is the average utility of decisions informed by knowledge of
the disturbance scenario $\tau$ (Figure \@ref(fig:figDisturbance)) and
priors from the national model:
\begin{equation}
EV_{current} = max_{\alpha}\sum_\psi[P(\psi|\tau)\eta_{\alpha,\psi}],
  (\#eq:EVcurrent)
\end{equation}
where $P(\psi|\tau)$ is the prior probability that the
true state of the population is $\psi$, and each population is assigned
a status $\alpha$ informed by the prior. The expected value under
certainty (i.e. the expected value of a decision when the true state of
the population is known) is $EV_{certainty}=1$. The difference between
$EV_{certainty}$ and $EV_{current}$ gives EVPI, a measure of the maximum
possible benefit that could theoretically be gained if the system were
perfectly known.

EVSI measures expected improvement from a particular monitoring strategy
as the difference between the expected value with and without additional
sample information: 
\begin{equation}
EVSI_{\tau} = E_x[max_{\alpha}\sum_\psi[P(\psi|x,\tau)\eta_{\alpha,\psi}]] - EV_{current},
  (\#eq:EVSI)
\end{equation}
where $x$ is the sample information from a particular
example trajectory. In this study, we treat simulated observations from
500 example trajectories for each disturbance and monitoring scenario
$\tau$ (Figure \@ref(fig:figDisturbance)) as samples from the
distribution of possible values of $x$, and assume each of these sample
results is equally probable. $P(\psi|x,\tau)$ is the posterior
probability that the true state of the population is $\psi$, and each
population is assigned a status $\alpha$ informed by the sample
information.

To aid in interpretation of results and to highlight the advantages of
selecting a metric that is directly relevant to the management objective
of stable or increasing populations, we also show how integrating local
monitoring data alters the distribution of differences between true
expected population growth rates and Bayesian posterior expected growth
rates.

# Results

Analysis of simulated example demographic trajectories confirms that our
Bayesian model can effectively combine local demographic data with
priors from a national demographic-disturbance relationships to reduce
uncertainty in boreal caribou demographic projections (Figures
\@ref(fig:examplePredictions) and \@ref(fig:figDiffs)). When local data are limited (e.g. 4 years, 15 collars per year) the model with informative priors (i.e. knowledge of national relationships) yields predictions that are consistent with knowledge of observed outcomes across the country, and the model with weakly informative priors does not (Figure \@ref(fig:examplePredictionsLowEffort)). When there is interannual variation increasing the duration of monitoring improves predictions of population growth rate substantially, and increasing the number of collars each year is less effective (Figure \@ref(fig:figDiffs)). Insufficiently informative priors can yield predictions that are both imprecise and inaccurate even in a case where there are 16 years of local monitoring data (Figure \@ref(fig:examplePredictions)), highlighting the benefits of integrating national relationships with these local data. Increasing the forecast horizon [@petchey_ecological_2015] does not substantially decrease the accuracy of expected population growth projections (compare 0, 5 and 20 year projections in
Figure \@ref(fig:figDiffs)) in these scenarios.

Examining the probability of correctly assessing population status (Figure \@ref(fig:figStatusErrors))
and the expected value of information (Figures \@ref(fig:figEVSI) and E.1) helps clarify how the need for monitoring varies with landscape condition. When anthropogenic disturbance is high (right panels of Figures
\@ref(fig:figStatusErrors), \@ref(fig:figEVSI), and E.1), projections of population status informed only by priors from the national demographic-disturbance relationships are accurate, and adding additional local monitoring data does not improve accuracy. At lower levels of disturbance, local monitoring can increase accuracy, but there are diminishing returns as the number of years of monitoring increases (Figures
\@ref(fig:figStatusErrors) and \@ref(fig:figEVSI)). Increasing accuracy is more difficult when interannual variation is high (Figures \@ref(fig:figStatusErrors) and \@ref(fig:figEVSI)). The shapes of the
relationships between monitoring effort and the probability of correct status assessment depend on how anthropogenic disturbance is changing over time, but there are diminishing benefits of increased monitoring
regardless of whether disturbance is increasing or decreasing (Supplement D). Acknowledging the possibility of composition survey errors reduces accuracy, but the reductions are not large for the set of parameters we investigate (Supplement F).

Averaging over the full set of example trajectories for each scenario (Figures \@ref(fig:figStatusErrors) and \@ref(fig:figEVSI)) obscures variation among them that is also important for monitoring decisions. For example, where disturbance is low but increasing, monitoring helps identify populations that are faring less well than expected now (top left of Figure \@ref(fig:figStatusErrorsLow)), and also helps identify populations that are likely to remain viable in the future when disturbance is higher (bottom right of Figure \@ref(fig:figStatusErrorsLow)). Status assessment accuracy is much higher when the 95% credible interval for expected population growth rate does not include one (`r round(100-erSums$probWrong[1],2)` %) than when it does (`r round(100-erSums$probWrong[2])` %), and this information could be used to guide monitoring. A more nuanced option is to recognize that the posterior probability of population viability ($P(\psi|x,\tau)$ in eq \@ref(eq:EVSI)) is an indicator of error probability (Figure \@ref(fig:figProbViability)). The probability of making an error also depends on the true (unknown) level of bias in the composition survey, but as the estimated posterior probability of viability approaches 0 or 1 accuracy increases to 100% even when the survey is biased (Figure \@ref(fig:figProbViability)). One could inform a decision about whether more information is needed for a particular population by selecting an error tolerance (e.g. the probability of correct status assessment should be >95 %) and using that threshold to distinguish cases where the posterior probability of viability is sufficiently certain (i.e. near enough to 0 or 1) from cases where more information is needed (Figure \@ref(fig:figProbViability)). A key advantage of accounting for composition bias uncertainty in the Bayesian model is that when the model indicates high confidence (i.e. the estimated probability of viability very close to 0 or 1) the answer is likely to be correct (high status assessment accuracy) even when the composition survey is biased (Figure \@ref(fig:figProbViability)).

# Discussion

## Bayesian integration of local demographic data & national demographic-disturbance relationships

The availability of boreal caribou demographic data varies widely across
Canada due to large variation in habitat condition, population
condition, available resources for monitoring, and remoteness. In a few
regions enough data have been collected over time to characterize local
demographic-disturbance relationships [@rudolph_demographic_2017;
@hervieux_widespread_2013], but most populations have only been sparsely
monitored for snapshot estimates of trend or abundance
[@johnson_science_2020]. As a result, in most regions it is necessary to
make assumptions informed by observed variation in demographic rates
with disturbance among populations and other relevant information about
impacts of disturbance on caribou. However, high observed variation in
demography where anthropogenic disturbance is low leads to high
uncertainty in projections informed by national demographic-disturbance
relationships alone [@stewart_climate-informed_2023;
@dyson_existing_2022]. In some cases, predictions of population growth
rate informed by national demographic-disturbance relationships appear
to disagree with estimates from local demographic data
[@eccc_scientific_2011; @dyson_existing_2022], but without formal
methods of accounting for uncertainty it can be difficult to distinguish
true discrepancies from stochastic variation.

We have proposed a simple Bayesian population model that helps address these challenges by
acknowledging uncertainty associated with local demographic data and
national demographic-disturbance relationships, and integrating these
two sources of information to reduce uncertainty in projections of the
impacts of changing landscape conditions. This Bayesian approach
represents an important improvement over less formally integrated
methods for estimating and projecting impacts of disturbance on
population caribou population viability [@eccc_scientific_2011;
@johnson_science_2020] because it reconciles and accounts for
uncertainties in local demographic data and in demographic-disturbance
relationships.

Our model assumes that increasing cumulative anthropogenic disturbance
causes a log-linear decrease in caribou demographic rates (equations 4,
5, 10 and 13), reflecting a scientific consensus that cumulative anthropogenic disturbance has a
negative impact on boreal caribou [@eccc_scientific_2011;
@johnson_science_2020; @gouvernement_du_quebec_revue_2021;
@rudolph_demographic_2017; @sorensen_determining_2008;
@hervieux_widespread_2013]. Local data can be integrated to estimate
local demographic-disturbance parameters that differ from the national
means, and in the absence of local data, appropriate priors lead to a
range of outcomes consistent with the observed range of variation across
the country [Figures 3 and 5 in @johnson_science_2020]. Others
[@moeller_integrating_2021; @mcnay_demographic_2022;
@eacker_web-based_2019; @dalgarno_bboutools_2024] have also used Bayesian models to estimate
caribou demographic rates from local monitoring data, but our addition
of prior information from national demographic-disturbance relationships
is novel, and important for reducing the risk that lack of local data
will be used as a reason to delay conservation action for highly
disturbed and poorly monitored populations [@bissonnette_controverses_2020; @boan_climate_2018]. Accounting for
demographic-disturbance relationships in a viability projection model
clarifies that no additional local information is needed to project population decline
where disturbance is high and actions to support recovery are not taken; the distribution of observed demographic rates from across the country (as summarized by the national analysis) is sufficient to predict problems for caribou on highly disturbed landscapes.

Interventions such as wolf culls, maternal penning, and restoration of
linear features are intended to improve survival and
reproduction of populations that are not otherwise self-sustaining
[@mcnay_demographic_2022; @serrouya_predicting_2020], and there is a
need to evaluate and compare effects of boreal caribou conservation
actions with careful monitoring. Our focus in this paper is on assessing
and projecting viability of populations in the absence of
these interventions, to clarify when additional information is needed to
reduce uncertainty about whether there is a need for action. The model we
propose does not include effects of management interventions, and
cannot be used to project outcomes for managed populations. Our intention is to reduce
uncertainty that can be used to delay conservation action; there is
nothing in our arguments, models or results that can or should be used
to discourage monitoring of the effectiveness of conservation
actions. Even when no additional monitoring is required for assessing
viability, there are a variety of reasons to continue
monitoring highly disturbed boreal caribou populations. The point we wish to emphasize is
that action is also required, and lack of monitoring should not be used
as a reason to delay conservation measures for highly disturbed
populations.

Variation in anthropogenic disturbance across the Canadian boreal region
is partially correlated with variation in climate and primary
productivity [@neilson_direct_2022] because forestry occurs in higher
productivity areas, and petroleum extraction activities are concentrated
in Western Canada. Effects of disturbance on vegetation and foodweb
dynamics vary with primary productivity [@neufeld_primary_2021], and
some have expressed concern about confounding of demographic-disturbance
relationships [@wilson_causal_2021]. One option for addressing these
concerns would be to develop regional demographic-disturbance
relationships for populations that are ecologically similar to one
another (e.g. caribou inhabiting forests that are sufficiently
productive for industrial forestry in Eastern Canada)
[@eccc_scientific_2024], and to use the priors from regional
demographic-disturbance relationships to inform demographic projections
within the region. However, while acknowledging and allowing for
regional variation, we also need to guard against risks that uncertainty
will be inflated and used as an excuse to delay action, a tactic that
has been used in this and other decision contexts
[@bissonnette_controverses_2020; @boan_climate_2018]. In order to be
informative and credible, regional demographic-disturbance analyses must
include a sufficient number of observations spanning a range from high
to low anthropogenic disturbance. Priors that lead to projections of
outcomes for caribou that fall outside the observed range of variation
across the country, characterized here (Figure 2) by 95% prediction
intervals from a national analysis (Johnson et al. 2020), are
implausible. In other words, if we have seen no examples of
self-sustaining caribou populations at a particularly high level of
disturbance, it is unreasonable to project or expect this outcome.

## Allowing for composition survey errors

Another novel aspect of this work is a term that acknowledges possible bias caused by misclassification of adults and missing calves in composition surveys [@ellington_divergent_2020;@decesare_estimating_2012]. Lacking information about the true magnitude of these errors, we examined the accuracy and precision of a Bayesian model that includes uncertainty about composition bias using examples with biased composition surveys. Acknowledging bias did decrease overall accuracy and precision, though these decreases were not large (Supplement E). Importantly, when the Bayesian model predictions were precise they were also accurate even if the true composition survey was biased (Figure 9). In other words, acknowledging composition bias uncertainty made predictions less confident, but in cases where predictions were confident they were also trustworthy.

A potential limitation of our analysis is that we assumed the same range of values for the probability of misclassifying adults and the probability of missing calves. Presuming that surveyed groups contain more cows than bulls, and that there is an equal probability of misclassifying cows as young bulls and vice versa, then misclassification errors cause high estimated recruitment. Missing calves causes low estimated recruitment, and together these errors tend to cancel one another out. An interesting extension of this work would be to investigate model performance over a broader range of assumptions about composition survey error rates, and to examine the implications of mismatches between the true distribution of errors and the unknown distribution in the Bayesian model. Better information about true composition survey error rates would also help assess their importance.

Alternatives to aerial composition surveys for estimating recruitment
include fecal pellet analysis [@mcfarlane_genetic_2018;
@mcfarlane_spatial_2021; @mcfarlane_identification_2022] and calf
collaring [@ellington_divergent_2020]. In an analysis of fecal pellets from 14 populations of boreal and southern mountain caribou, Jones et al. [-@jones_novel_2023] found that sampled
groups contained more females than males, and that cows tend to stay
with their male and female offspring, while bulls do not. These results are
broadly consistent with our assumption that composition survey groups
include cows and younger animals of both sexes that can be difficult
to classify visually. Documenting animals that are difficult to classify
as “unknown” can help analysts account for classification uncertainties [@ellington_divergent_2020;@dalgarno_bboutools_2024], as can comparing classification
results from multiple independent observers. The probability of missing
calves likely increases with vegetation cover, and is more difficult to quantify.

Each survey method has costs and benefits, and the reality is that monitoring everything everywhere all the time in great detail is neither feasible nor necessary. We hope that
adding a bias term to the simple recruitment-mortality model that is used for many herd viability assessments [@hatter_moose_1991;@hatter_revisiting_2020;@decesare_estimating_2012;
@serrouya_comparing_2017; @eacker_web-based_2019; @eccc_scientific_2011;
@johnson_science_2020; @dalgarno_bboutools_2024] will enable and encourage practitioners to examine the sensitivity of their conclusions to plausible composition survey error rates, and help clarify thinking about where, when, and why additional information is needed.

## Limitations and Future Possibilities

The main objectives of Johnson et al's [-@johnson_science_2020;
@eccc_scientific_2011] national analysis were to inform a national
disturbance policy that applies to all ranges except one in northern Saskatchewan (SK1) that is subject to very low anthropogenic disturbance and high fire disturbance, and to inform
a different policy for SK1 [@eccc_scientific_2011; @eccc_boreal_2019]. A
hierarchical analysis [@leasure_hierarchical_2019] would be better
suited for estimating variation in demographic model parameters among
populations, and for distinguishing that from variation among years. We have not reanalyzed the national data using different methods because the data are owned by many jurisdictions, and obtaining permission from all of these jurisdictions would be difficult and time-consuming. To move ahead with available published information we constrained our example trajectory simulations by assuming that populations remain in their quantile, so populations that are lower than or higher than average remain lower or higher than average over time.
Note that the Bayesian model is not similarly constrained, and can accommodate
evidence of a shift from unexpectedly low to unexpectedly high
demographic rates over time, or vice versa. Switching from low to high
each year is unlikely, but there is evidence of variation in the slope
of demographic-disturbance relationships among populations
[@rudolph_demographic_2017], so it is important that the Bayesian model
allows for a greater range of possibilities than our more constrained
simulated example trajectories. Although the simulated trajectories do
not include all plausible possibilities, we believe they are sufficient
to demonstrate how a Bayesian model can be used to integrate local data
with national priors, and to show that monitoring requirements depend on
landscape condition.

For this analysis we used a simple survival analysis method that does not account for variation in survival rate or the number of collared animals throughout the year. More sophisticated methods should be used for analysis of real survival data, but our implementation of a two-step non-parametric survival analysis method [following @eacker_web-based_2019] yielded slightly biased survival estimates. Dalgarno et al.’s [-@dalgarno_bboutools_2024] method may be a better alternative. More generally, the proof-of-concept model presented in this paper is adequate for making general points about the integration of national demographic-disturbance relationships with local data, and for accounting for composition survey errors. However, it is a research prototype, not a fully developed analysis tool. The more fully developed and documented *bboutools* R package [@dalgarno_bboutools_2024] is likely a better starting point for development of tools to enable local decision makers to inform decisions with their existing data. Possible next steps are to extend that package to include priors from demographic-disturbance relationships and composition survey errors.

Aspatial demographic models generally assume that observed animals are a
random and representative sample of a fully mixed population in which
survival and recruitment rates apply equally to all animals. However,
metrics derived from fecal pellet analysis such as reproductive success
[@mcfarlane_genetic_2018; @mcfarlane_spatial_2021;
@mcfarlane_identification_2022], age structure and pregnancy rates
[@flasko_fecal_2017; @mcfarlane_identification_2022], and
mother-offspring co-occurence [@jones_novel_2023] suggest high variation
in fitness among individuals [@mcfarlane_genetic_2018] and spatial
structure within populations [@jones_novel_2023;
@mcfarlane_spatial_2021; @mcfarlane_identification_2022]. Analyses of
movement data also show spatial variation in risks and benefits for
caribou and variation in habitat preferences and space use among
individuals [e.g. @avgar_spaceuse_2015; @decesare_linking_2014]. It is possible to integrate complexities such as variation in demographic rates
among individuals [@armstrong_incorporating_2021], population size
information from aerial or fecal DNA surveys, and variation in demographic rates among age classes [@moeller_integrating_2021; @mcnay_demographic_2022] into Bayesian population models. However, in many contexts there are not enough to data to estimate more complex model parameters, and simple models remain relevant.

Our proof-of-concept analysis shows the value of integrating national demographic-disturbance relationships with local data, accounting for composition survey errors, and assessing the value of local monitoring
information. We applied the framework proposed by Dunham et al. (2023) to show that
no additional local monitoring information is needed to predict whether
boreal caribou populations in highly disturbed landscapes are viable in the absence of interventions such as predator control or maternal penning; self-sustainability is unlikely. There is more variation in caribou demographic rates
among study areas with low to moderate disturbance, and thus more need
for local data to inform demographic projections. In these places, our
analysis suggests that there may be diminishing returns from
increasing monitoring effort, but more information on monitoring costs,
objectives, and existing local data would be required to assess the
adequacy and cost-effectiveness of monitoring scenarios in particular
circumstances. Monitoring costs vary substantially with technologies
used, the remoteness of herds, and the difficulty of collaring animals,
which in turn depends on the number of currently collared animals, the
population size, and vegetation across the landscape
[@nbckc_boreal_2021]. It is important to recognize that projecting
population viability is only one possible objective for monitoring
[@nbckc_boreal_2019], and that the impacts of monitoring on threatened
animals should be considered [@nbckc_boreal_2019]. Since monitoring
costs, options and goals are context specific, we believe the most
effective way forward will be to use this proof-of-concept analysis, existing local data, and the more fully developed *bboutools* R package [@dalgarno_bboutools_2024] as a starting point for working with interested decision makers to develop monitoring decision support tools that meet real local needs, recognizing that those needs will vary [@nbckc_nbckc_2021;@nbckc_nbckc_2022].

# Supplementary material
Supplement A: Definitions of variables and parameters in the Bayesian population model
Supplement B: Model of bias in recruitment estimates from calf:cow surveys
Supplement C: Calibration of priors
Supplement D: Alternative disturbance and monitoring scenarios
Supplement E: Expected value of perfect information
Supplement F: Sensitivity to unknown bias in recruitment estimates from calf:cow surveys

# Data statement
Code required to create the data, run analyses and reproduce figures in this paper is available at https://github.com/LandSciTech/Caribou-Demographic-Projection-Paper, and in the caribouMetrics R package available on GitHub at https://github.com/LandSciTech/caribouMetrics/tree/EI-paper-submission.

# Acknowledgements
This work was informed by a variety of discussions with more people than we can name, but here are a few. Members of the Monitoring Working Group of the National Boreal Caribou Knowledge Consortium and the Scientific Assessment of Federal and Provincial Frameworks for the Conservation of Boreal Caribou in Ontario helped us understand context. Tal Avgar and Micheline Manseau encouraged us to think about composition survey errors. Comments from Joseph Bennett and Daniel Eacker on early versions of this work motivated a variety of improvements. Note that our acknowledgement of these contributions does not imply endorsement; we also acknowledge disagreement on some points, and that all remaining errors are our own.

# Author contributions
**Josie Hughes:** Conceptualization, Methodology, Formal analysis, Investigation,
Supervision, Software, Visualization, Writing – original draft, Writing – review and editing
**Sarah Endicott:** Resources, Software, Writing – review and editing
**Anna M Calvert:** Conceptualization, Methodology, Writing – original draft, Writing – review and editing
**Cheryl A Johnson:** Conceptualization, Methodology, Writing – review and editing

<!-- The following line inserts a page break -->

\newpage

# References

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files): -->

::: {#refs}
:::

# Figures

```{r figDisturbance, fig.cap=".", cache=TRUE}
knitr::include_graphics(here::here("figs/forPaper/distScnscmult6ay1aSf1.png"))
file.rename(here::here("analysis/paper/figs_submit/distScnscmult6ay1aSf1.pdf"), 
            here::here("analysis/paper/figs_submit/Figure1_distScnscmult6ay1aSf1.pdf"))
```

```{r exampleTrajectories, fig.cap=".", cache=TRUE}
knitr::include_graphics(here::here("analysis/paper/figs/DemographicRates.png"))
file.rename(here::here("analysis/paper/figs_submit/DemographicRates.pdf"), 
            here::here("analysis/paper/figs_submit/Figure2_DemographicRates.pdf"))
```

```{r examplePredictions, fig.cap=".", cache=TRUE}
knitr::include_graphics(here::here("analysis/paper/figs/bayesianModelExamples.png"))
file.rename(here::here("analysis/paper/figs_submit/bayesianModelExamples.pdf"), 
            here::here("analysis/paper/figs_submit/Figure3_bayesianModelExamples.pdf"))
```

```{r examplePredictionsLowEffort, fig.cap=".", cache=TRUE}
knitr::include_graphics(here::here("analysis/paper/figs/bayesianModelExamplesLowEffort.png"))
file.rename(here::here("analysis/paper/figs_submit/bayesianModelExamplesLowEffort.pdf"), 
            here::here("analysis/paper/figs_submit/Figure4_bayesianModelExamplesLowEffort.pdf"))
```

```{r figDiffs, fig.cap=".", cache=TRUE}
knitr::include_graphics(here::here("figs/forPaper/diffslowcmult6ay1aSf1.png"))
file.rename(here::here("analysis/paper/figs_submit/diffslowcmult6ay1aSf1.pdf"), 
            here::here("analysis/paper/figs_submit/Figure5_diffslowcmult6ay1aSf1.pdf"))
```

```{r figStatusErrors, fig.cap=".", cache=TRUE}
knitr::include_graphics(here::here("figs/forPaper/powercmult6ay1aSf1.png"))
file.rename(here::here("analysis/paper/figs_submit/powercmult6ay1aSf1.pdf"), 
            here::here("analysis/paper/figs_submit/Figure6_powercmult6ay1aSf1.pdf"))
```

```{r figEVSI, fig.cap=".", cache=TRUE}
knitr::include_graphics(here::here("figs/forPaper/EVSIcmult6ay1aSf1.png"))
file.rename(here::here("analysis/paper/figs_submit/EVSIcmult6ay1aSf1.pdf"), 
            here::here("analysis/paper/figs_submit/Figure7_EVSIcmult6ay1aSf1.pdf"))
```

```{r figStatusErrorsLow, fig.cap=".", cache=TRUE}
knitr::include_graphics(here::here("figs/forPaper/powerLamcmult6ay1aSf1.png"))
file.rename(here::here("analysis/paper/figs_submit/powerLamcmult6ay1aSf1.pdf"), 
            here::here("analysis/paper/figs_submit/Figure8_powerLamcmult6ay1aSf1.pdf"))
```

```{r figProbViability, fig.cap=".", cache=TRUE}
knitr::include_graphics(here::here("figs/forPaper/probViabilitycmult6ay1aSf1.png"))
file.rename(here::here("analysis/paper/figs_submit/powerLamcmult6ay1aSf1.pdf"), 
            here::here("analysis/paper/figs_submit/Figure9_probViabilitycmult6ay1aSf1.pdf"))
```

## Figure Captions
Figure 1: Anthropogenic disturbance and monitoring scenarios. Anthropogenic disturbance increases by 1% per year, and the amount of disturbance at the beginning of the projection period varies from low (0%), through low-med (20%) and med-high (40%) to high (60%). These simple hypothetical scenarios are intended to show how the need for local monitoring to reduce uncertainty varies with landscape state, and not intended to reflect real landscape change. In all scenarios, the projection period begins in 2024, and the monitoring period occurs prior to the projection period.

Figure 2: Examples of simulated population trajectories from a model informed only by national disturbance-demographic relationships. To highlight how variation among trajectories persists as disturbance changes, we have reduced interannual variation by 75% in these examples ($v_R=0.115,v_S=0.02175$). Orange dots show means of 500 example trajectories. For reference, black lines and and shaded bands show means and 95% predictive intervals from Johnson et al's (2020) regression models.

Figure 3: Prior (a) and posterior predictions from the Bayesian model for population trajectories with informative (a,b) and weakly informative (c) priors given lower interannual variation ($v_R=0.23,v_S=0.435$) and long-term monitoring. In this example scenario anthropogenic disturbance increases by 2% per year from 0 to 100%. The population is monitored for 16 years with 30 collars per year, and 6 cows per collared cow in the composition surveys. The prior means and 95% predictive intervals (a, orange lines and shading) are similar to the means and ranges between the 2.5% and 97.5% quantiles of 3000 simulated survival and recruitment trajectories from the national model (a, blue lines and shading). Local population data (open triangles) reduces uncertainty about demographic rates (b), though that is not needed to predict viability (population growth <= 0.99) of highly disturbed populations (a). With weakly informative priors the model does not provide accurate or precise predictions of future viability after 16 years of monitoring in a case where anthropogenic disturbance is increasing to 90% (c).

Figure 4: Posterior predictions from the Bayesian model for population trajectories with informative (a) and weakly informative (b) priors given higher interannual variation ($v_R=0.46,v_S=0.087$) and limited monitoring (4 years, 15 collars per year, 3 cows per collared cow in composition surveys). See Figure 3 for disturbance trends and other details. The limited local data (open triangles) reduces uncertainty (orange shading) about the observed current state of the population, but is not sufficient to reduce future uncertainty. Given limited data the model with informative priors (a) yields predictions (orange bands) that are consistent with knowledge of observed outcomes across the country (blue shading), and the model with weakly informative priors (b) does not.

Figure 5: The distribution of differences between true expected population growth rates and Bayesian posterior expected growth rates varies with the amount of anthropogenic disturbance (columns), the number of years projected (rows), and monitoring effort (see Figure \@ref(fig:figDisturbance) for scenarios). Differences decrease as the number of years of monitoring and the number of collars increase. In these examples the composition survey cow multiplier $w$ is 6, collars are renewed each year, and interannual variation is lower ($v_R=0.23,v_S=0.0435$).

Figure 6: The probability of a correct status assessment varies with the amount of anthropogenic disturbance (columns), the number of years projected (rows), the amount of interannual variation (high: $v_R=0.46,v_S=0.087$, low: $v_R=0.23,v_S=0.0435$, none), and monitoring effort (see Figure \@ref(fig:figDisturbance) for scenarios). When disturbance is high the probability of correct status assessment is high even with one year of monitoring. Where disturbance is low or moderate, increasing the number of years of monitoring increases accuracy, but there are diminishing returns on additional monitoring as the number of years increases. In these examples the composition cow multiplier $w$ is 6 and collars are renewed each year; varying these parameters does not alter qualitative conclusions (Supplements D and E).

Figure 7: The expected value of sample information (EVSI) varies with the amount of anthropogenic disturbance (columns), the number of years projected (rows), the amount of interannual variation (high: $v_R=0.46,v_S=0.087$, low: $v_R=0.23,v_S=0.0435$, none), and monitoring effort. When disturbance is high EVSI is low, indicating that local monitoring does not change predicted status. Where disturbance is lower, EVSI increases with the number of years of monitoring and the number of collars, but there are diminishing returns on additional monitoring as the number of years increases. See Figure \@ref(fig:figStatusErrors) for additional scenario details.

Figure 8: When anthropogenic disturbance is low and expected to increase (low scenario in Figure \@ref(fig:figDisturbance)) the probability of a correct status assessment varies with the true growth rate of the population (columns), the number of years projected (rows), the amount of interannual variation (high: $v_R=0.46,v_S=0.087$, low: $v_R=0.23,v_S=0.0435$, none), and monitoring effort (see Figure \@ref(fig:figDisturbance) for scenarios). Increasing the number of years of monitoring helps identify populations with currently declining populations (top left), and to identify populations with higher than expected growth rates that fare better as disturbance increases (lower right). Increasing the number of collars more effectively increases near-term accuracy when interannual variation is low (top left). If the goal is long-term projection (bottom row) there is little advantage to increasing the number of collars. Increasing the accuracy of status assessment for populations with growth rates close to one is difficult (middle columns). In these examples the composition cow multiplier $w$ is 6 and collars are renewed each year.

Figure 9: The posterior probability that the population is viable ($P(\psi|x,\tau)$ in eq \@ref(eq:EVSI)) predicts the probability of a correct status assessment (estimated from the 114,000 example cases in Figures 5 to 7 with a generalized additive model). The relationship between accuracy and the probability of viability depends on true composition survey bias (colours), but regardless of bias accuracy is high when the probability of viability is very near to 0 or 1. Cases in which there is no need for additional monitoring if one is willing to accept a 5% error rate fall above the dashed line.

### Colophon

This report was generated on `r Sys.time()` using the following
computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
#if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
#if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```